### 2024-05-24 scripts/first_pass_adapting_K23_analysis_20231031.Rmd
get scripts in there

---
title: "MS Prelim Data K23"
author: "Erica Baller"
date: "2/25/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(stringr)
#functions
source("~/BBL/msanxiety/scripts/ms_functions.R")
```

## MS anxiety analyses

## R Markdown

The first chunk here is reading in appropriate data frames and preprocessing. Two big data frames are read in. The first is the demographic, drug, lab info for the MS patients (from the DAC pull). The second is the data frame of fascicle proportion affected for people with good mimosa (# volume affected/total volume of fascicle). The most important preprocessing occurs in a series of mutate commands. These essentially add new columns to the data frame that can later be used. The thrust of these mutate commands is to define who has depression and who does not.

Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion. Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses. For these preliminary analysis, our inclusion/exclusion are more stringent. Of note, the graph below is irrespective of whether they had good/bad MIMoSA. As such, the MIMoSA subset is smaller. It does however highlight the

<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>

Inclusion criteria below:

Inclusion MS Criteria (dataframe read in here has already been filtered down using the exclusions below previously)
  1) Scanned under MS protocol (42K)
  2) MS Diagnosis (17K)
  3) Seen by MS doctor (16K)

Inclusion criteria for Anxious Group
  1) Have an ICD10 F3* code (in our sample, people have F41.1; 41.3; 41.8; 41.9)
        (may use this later)- https://www.nami.org/About-Mental-Illness/Treatments/Mental-Health-Medications

Inclusion criteria for Healthy Group
  1) No F* diagnoses
  2) At least 1 PHQ2 or PHQ9 with a documented 0
  3) Free of psychiatric medications per NAMI website

Exclusions
  If you received the MS protocol or had a diagnosis of MS but never actually saw an MS provider, we cannot be sure that the diagnosis of MS is truly valid, as it is hard to diagnose anyway.
  If you did not have a diagnosis of depression and never completed a PHQ2 or 9, and had no history of medications, we could not safely confirm nor deny the presence of depression. These subjects were excluded from future analyses. We could consider adding these to the healthy group maybe, but not too inclined to do that right now.

Imaging exclusion
  We are currently only using people who had good MIMOSA (75 or 100). That severely limits our numbers, but that is okay! When we re-QC, we will probably get a lot more into this group.

Final N demographics 
  Anxious: 
  Super clean healthy: 604 (148 unique)

(unique, with good MIMoSA):

```{r read_in_df_and_subset}
homedir <- "/Users/eballer/BBL/msanxiety/"
columns_to_make_integer <- c("ACCESSION_NUM", "PAT_AGE_AT_EXAM","MRI_ENC_AGE","CURRENT_AGE","hemoglobin", "WBC","RBC","B12","FOLATE", "TSH", "RPR","CSFAPPEAR1","CSFAPPEAR4","CSFCOLOR1",       "CSFCOLOR4","CSFTUBE","CSFTUBE4","PHQ.2","PHQ.9") #removed "VitD", not in new pull for some reason

#open Rds
data_2023_pull <- readRDS(paste0(homedir, "/data/data_2023_pull_with_curated_icd10_codes.rds"))

#Get med info in
#read in BCP info, names of meds, est and prog and doses
bcp <- read.csv("/Users/eballer/BBL/medication_data/BCP_est_and_prog_levels_and_doses_all_brands.csv", sep = ",", header = TRUE)
bcp$bcp_name <- toupper(bcp$bcp_name)

#read in ms med info, names of meds, mechanism, and whether or not dmt
msmeds <- read.csv("/Users/eballer/BBL/medication_data/ms_medications_brand_and_generic.csv", sep = ",", header = TRUE)
msmeds$Medication <- toupper(msmeds$Medication)

#depression/psych meds from NAMI website 
psych_meds <- read.csv("/Users/eballer/BBL/medication_data/nami_psych_meds_antidepressants.csv", sep = ",", header = TRUE)
psych_meds$Medication.Name <- toupper(psych_meds$Medication.Name)

#anxiety meds from chatgpt (cross referenced with psych meds)
anxiety_meds <- read.csv("/Users/eballer/BBL/medication_data/antianxiety_medications.csv", sep = ",", header = TRUE)
anxiety_meds$Medication <- toupper(anxiety_meds$Medication)
anxiety_meds_no_beta_blocker <- anxiety_meds %>% filter(Class != "Beta_blocker")

##########################
### some preprocessing ###
##########################
#we need to keep scans from people seen by an MS provider (n = 17067) who have an MS diagnostic code (n=16,830)
#we make a bunch of columns from character to integer, binarize sex and race, and put date into a nice format YYYYMMDD, the %m%d%y indicates what it started out as

#goal is to keep people who were seen by 
 data_empi_acc_f_phq <-data_2023_pull %>% 
   mutate(across(.cols = columns_to_make_integer, .fns = as.integer)) %>%
   mutate(sex_binarized = ifelse(SEX == "MALE", 1, 2)) %>%
   mutate(osex = ordered(sex_binarized,levels = c(1,2), labels = c("Male","Female"))) %>%
   mutate(race_binarized = ifelse(RACE == "WHITE", 1, 2)) %>%
   mutate(orace = ordered(race_binarized,levels = c(1,2), labels = c("White","Non-white"))) %>%
   mutate(EXAM_DATE = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% 
   mutate(EXAM_DATE = gsub(EXAM_DATE, pattern = "-", replacement = "")) %>%
   mutate(BEGIN_EXAM_DTTM = as.Date(BEGIN_EXAM_DTTM, format = "%m/%d/%y")) %>% #make sure to actually save it as date object for later processing
   mutate(EMPI = as.factor(EMPI)) %>%
   mutate(number_of_dx = ifelse(!is.na(ICD10), (str_count(ICD10, ";") + 1), 0)) %>%
   mutate(number_of_dx_extended = ifelse(!is.na(all_icd10_dx), str_count(all_icd10_dx, ";") + 1, 0)) %>%
   mutate(number_of_psychiatric_comorbidities_extended = ifelse(!is.na(all_icd10_dx), str_count(all_icd10_dx, "F"), 0)) %>%
   mutate(number_of_anxiety_comorbidities_extended = ifelse(!is.na(all_icd10_dx), str_count(all_icd10_dx, "F4"), 0)) %>%
   mutate(number_of_depression_comorbidities_extended = ifelse(!is.na(all_icd10_dx), str_count(all_icd10_dx, "F3"), 0)) %>%
   mutate(number_of_medications = ifelse(!is.na(Medication), str_count(Medication, ";") + 1, 0)) %>%
   mutate(On.Bcp = ifelse(str_detect(Medication, paste(bcp$bcp_name, collapse = "|")), 1, 0)) %>%
   mutate(On.Psych.Meds = ifelse(str_detect(Medication, paste(psych_meds$Medication, collapse = "|")), 1, 0)) %>%
   mutate(On.Antidepressants = ifelse(str_detect(Medication, paste(psych_meds$Medication[psych_meds$Antidepressant == "t"], collapse = "|")), 1, 0)) %>%
   mutate(On.Anxiolytics = ifelse(str_detect(Medication, paste(anxiety_meds$Medication, collapse = "|")), 1, 0)) %>%
   mutate(On.Anxiolytics_no_beta_blocker = ifelse(str_detect(Medication, paste(anxiety_meds_no_beta_blocker$Medication, collapse = "|")), 1, 0)) %>%
   mutate(On.Benzos = ifelse(str_detect(Medication, paste(anxiety_meds$Medication[anxiety_meds$Class == "Benzodiazepine"], collapse = "|")), 1, 0)) %>%
   mutate(On.Anti_cd20 = ifelse(str_detect(Medication, paste(msmeds$Medication[msmeds$Mechanism == "anti-cd20"], collapse = "|")), 1, 0)) %>%
   mutate(On.Interferon = ifelse(str_detect(Medication, paste(msmeds$Medication[msmeds$Mechanism == "Interferon"], collapse = "|")), 1, 0)) %>%
   mutate(On.Steroids = ifelse(str_detect(Medication, paste(msmeds$Medication[msmeds$Mechanism == "Steroid"], collapse = "|")), 1, 0)) %>%
   mutate(On.Dmt = ifelse(str_detect(Medication, paste(msmeds$Medication[msmeds$DMT == "y"], collapse = "|")), 1, 0)) %>%
   mutate(Has.PHQ2 = ifelse(PHQ.2 != "NULL", 1, 0)) %>%
   mutate(Has.PHQ9 = ifelse(PHQ.9 != "NULL", 1, 0)) %>%
   mutate(Has.depdx = ifelse(grepl("F3", all_icd10_dx), 1, 0)) %>%
   mutate(Has.anxietydx = ifelse(grepl("F4", all_icd10_dx), 1, 0)) %>%
   mutate(Has.depANDanx.dx = ifelse(Has.depdx & Has.anxietydx, 1, 0)) %>%
   mutate(PHQ.2_modsev_dep_sxs = ifelse(Has.PHQ2 & PHQ.2 >= 3, 1, 0)) %>%
   mutate(PHQ.9_modsev_dep_sxs = ifelse(Has.PHQ9 & PHQ.9 >= 10, 1, 0)) %>%
   mutate(PHQ.2_mild_dep_sxs = ifelse(Has.PHQ2 & PHQ.2 > 0 & PHQ.2 < 3, 1, 0)) %>%
   mutate(PHQ.9_mild_dep_sxs = ifelse(Has.PHQ9 & PHQ.9 > 0 & PHQ.9 < 10, 1, 0)) %>%
   mutate(PHQ.2_zero = ifelse(Has.PHQ2 & PHQ.2 == 0, 1, 0)) %>%
   mutate(PHQ.9_zero = ifelse(Has.PHQ9 & PHQ.9 == 0, 1, 0)) %>%
   mutate(dep_by_dx_phq = ifelse((Has.depdx | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(dep_by_dx_phq_antidep = ifelse((Has.depdx | On.Antidepressants | PHQ.2_modsev_dep_sxs | PHQ.9_modsev_dep_sxs),1,0)) %>%
   mutate(true_healthy = ifelse((!Has.depdx & !On.Psych.Meds & (PHQ.2_zero | PHQ.9_zero)), 1, 0)) %>%
   mutate(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds = ifelse(dep_by_dx_phq_antidep | true_healthy, 1, 0)) %>%
   rowwise(ACCESSION_NUM) %>%
   mutate(depGroupVar =
            sum(c(dep_by_dx_phq_meds_healthy_phq0_no_psych_meds,dep_by_dx_phq_antidep))) %>%#you get an extra point if you are in the depression group, so the end result is dep = 2, healthy = 1, 0 for exclude
   ungroup() #n=3,737 unique people, n= 16,830 total

##########################################################
#             Read in Fascicle info             #
##########################################################

fascicle_proportions <- read.csv(paste0(homedir, "/results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])

write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)

########
#fascicle volumes in depression network
########
fascicle_volumes_dep <- read.csv(paste0(homedir, "results/streamline_volume_within_dep_network.csv"), header = T, sep = ",")

#add column to indicate whether does not meet criteria for inclusion in dep mask (0), any overlapping voxels (1), or >10% overlapping voxels (2)
fascicle_volumes_dep_network <- fascicle_volumes_dep %>%
  replace(is.na(.), 0) %>%
  mutate(inDepMask=case_when(prop_in_mask == 0 ~ 0, (prop_in_mask > 0 & prop_in_mask < 0.1) ~ 1, prop_in_mask >= 0.1 ~ 2))

write.table(fascicle_volumes_dep_network, paste0(homedir,  "/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_grouping.csv"), sep = ",", col.names=T)#, row.names = F, quote = F, col.names = T)

fascicle_names_nondep_network<- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 0)]

fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask > 0)]

#fascicle_names_dep_network_all <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 1)]
#write.table(fascicle_names_dep_network_all,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_all.csv", row.names = F, quote = F, col.names = F)

fascicle_names_dep_network_10_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$inDepMask == 2)] #n=15
#write.table(fascicle_names_dep_network_10_percent,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names_dep_network_10_percent.csv", row.names = F, quote = F, col.names = F)

fascicle_names_dep_network_5_percent <- fascicle_volumes_dep_network$fascicle[which(fascicle_volumes_dep_network$prop_in_mask >= 0.05)] #n=24

#fascicle_mapping, returns numerical mapping as well as names of tracts, and whether in or out of dep network
fascicle_bundle_mapping <- get_fascicle_bundle_mapping()

##########################################################
#             Merge with data_empi              #
##########################################################
df_demo_and_fascicles <- merge(data_empi_acc_f_phq, fascicle_proportions, by = c("EMPI", "EXAM_DATE")) #n = 2962

##########################################################
#                 Histos                        #
##########################################################
##### Histograms for PHQ2/9 healthy and depressed

hist(df_demo_and_fascicles$PHQ.2[which(!is.na(df_demo_and_fascicles$PHQ.2))], main = paste0("PHQ-2 In Good Mimosa Group : n = ", length(which(!is.na(df_demo_and_fascicles$PHQ.2)))), xlab = "PHQ-2", breaks = 8)

hist(df_demo_and_fascicles$PHQ.9[which(!is.na(df_demo_and_fascicles$PHQ.9))], main = paste0("PHQ-9 In Good Mimosa Group : n = ", length(which(!is.na(df_demo_and_fascicles$PHQ.9)))), xlab = "PHQ-9", breaks = 8)

phq2_and_dep <- df_demo_and_fascicles %>% filter(depGroupVar == 2) %>% drop_na(PHQ.2) 
phq9_and_dep <- df_demo_and_fascicles %>% filter(depGroupVar == 2) %>% drop_na(PHQ.9) 
phq2_and_healthy <- df_demo_and_fascicles %>% filter(Has.depdx == 0) %>% drop_na(PHQ.2) 
phq9_and_healthy <- df_demo_and_fascicles %>% filter(Has.depdx == 0) %>% drop_na(PHQ.9)

hist(phq2_and_dep$PHQ.2, main = paste0("PHQ-2/Depressed In Good Mimosa Group : n = ", dim(phq2_and_dep)[1]), xlab = "PHQ-2", breaks = 8)

hist(phq9_and_dep$PHQ.9, main = paste0("PHQ-9/Depressed In Good Mimosa Group : n = ", dim(phq9_and_dep)[1]), xlab = "PHQ-9", breaks = 8)

hist(phq2_and_healthy$PHQ.2, main = paste0("PHQ-2/Healthy In Good Mimosa Group : n = ", dim(phq2_and_healthy)[1]), xlab = "PHQ-2", breaks = 8)
hist(phq9_and_healthy$PHQ.9, main = paste0("PHQ-9/Healthy In Good Mimosa Group : n = ", dim(phq9_and_healthy)[1]), xlab = "PHQ-9", breaks = 8)

######## Histograms for depression network
#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
hist(x=fascicle_volumes_dep$prop_in_mask, main = "Histogram Affected Streamlines in Dep Net", xlab = "Proportion Streamlines per Fascicle Affected", breaks = 20)

```

## Depression versus healthy analysis

```{r demographics}

#this will display a demographics table comparing age, race, sex between healthy and depressed groups, using the first presentation to treatment

make_demographics_table_ms(data_frame = data_empi_acc_f_phq)

df_unique_empi <- data_empi_acc_f_phq %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

make_demographics_table_ms(data_frame = df_unique_empi)

df_just_healthy_and_depressed <- df_unique_empi %>% filter(depGroupVar != 0)

make_demographics_table_ms(data_frame = df_just_healthy_and_depressed)

#just people with good mimosa
df_good_mimosa <- df_demo_and_fascicles %>% filter(depGroupVar != 0)
make_demographics_table_ms(data_frame = df_good_mimosa)

df_good_mimosa_unique <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique)
```

```{r proof_of_concept}

#this section creates histograms looking at the proportion of volume lost in each fascicle across subjects
#we recapitulate earlier work and show that higher volume loss with our method reflects regions previously known to have higher volume loss
#These findings suggest that the streamline filtering approach could be a reasonable approach for this analysis

#have to pull out just the pieces I'm interested in
just_dep_and_empi <- subset(data_empi_acc_f_phq, select = c("EMPI", "depGroupVar"))
added_dep_to_fascicles <- merge(just_dep_and_empi, fascicle_proportions, by = c("EMPI"))
melted_df <- melt(added_dep_to_fascicles, id.vars = c("EMPI", "EXAM_DATE", "depGroupVar"))

#exclude regions where there was no lesion at all
lesioned <- subset(melted_df, value > 0)

#only include people in healthy or depressed group
lesioned_dx <- subset(lesioned, depGroupVar != 0)

#separate out healthy and depressed
healthy <- subset(lesioned, depGroupVar == 1)
depressed <- subset(lesioned, depGroupVar == 2)

q<-ggplot(lesioned, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r<-ggplot(lesioned_dx, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r_color<-ggplot(lesioned_dx, aes(x=value, color=factor(depGroupVar))) + geom_histogram() + facet_wrap(~variable)
x<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
y<-ggplot(healthy, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
z<-ggplot(depressed, aes(x=value)) + geom_histogram() + facet_wrap(~variable)

#print(q)
#print(r)
print(r_color)
#print(x)
#print(y)
#print(z)
```

```{r age_effects_fascicles_unique}
########### age effect whole group

#isolate out only the depressed group and healthy group
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#### age histo ###
hist(dep_and_healthy_groups_for_ICD_analysis$PAT_AGE_AT_EXAM,
main = paste0("Age In Good Mimosa Group : n = ", dim(dep_and_healthy_groups_for_ICD_analysis)[1]), xlab = "Age", breaks = 10)

#lm
fascicle_age_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_age_lm) <- fascicle_names

#anova
fascicle_age_anova <- lapply(fascicle_age_lm, anova)

#fdr corrected
fascicle_age_anova_fdr <- fdr_anova_generic(fascicle_age_anova, 1)
print(fascicle_age_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_age_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#visreg
#sapply(fascicle_age_lm,visreg)
visreg(fascicle_age_lm$AF_R,ylab = "Proportion of Affected Streamlines", xlab = "Age in Years",
       line=list(col="red"),
       points=list(col="black"))

```

```{r sex_effects_fascicles_unique}

### really not interesting at all. 3 areas that are uncorrected: SCP, CPT_O_R, OR_R

#isolate out only the depressed group and healthy group
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#### sex histo ###
hist(dep_and_healthy_groups_for_ICD_analysis$sex_binarized,
main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1), "/", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)),  xlab = "Sex", breaks = 2)

#lm
fascicle_sex_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_sex_lm) <- fascicle_names

#anova
fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)

#fdr corrected
fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
print(fascicle_sex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_sex_anova_p <- sapply(fascicle_sex_anova, function(v) v$"Pr(>F)"[1])
fascicle_sex_anova_p05_unc <- as.data.frame(fascicle_sex_anova_p[fascicle_sex_anova_p < 0.05])
print(fascicle_sex_anova_p05_unc)

```

```{r depression_unique_all_fascicles_together_in_dep_networks}

#isolate out only the depressed group and healthy group, make a column that sums up the total prop lost)
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
  mutate(total_fascicle_prop_lost_indepnet_10_percent =
           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
   mutate(total_fascicle_prop_lost_indepnet_5_percent =
           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
  mutate(total_fascicle_prop_lost_nondep_network =
           rowMeans(across(fascicle_names_nondep_network))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

dep_vs_healthy <- lm(total_fascicle_prop_lost ~ depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis)
summary(dep_vs_healthy) #p=0.494, cohen d = 0.2

#power analysis to predict sample required for K, u = num df, v = denom df, f2=0.01 (cohen d = 0.02/2) 
#prop in each fascicle predicting dep (87), number of predictors (age, sex, ef, ess?) = 4, u = 87-4 = 83
pwr.f2.test(u=83, f2 = 0.02, sig.level = 0.5, power = 0.8)

#test whether proportion lost overall different between depressed and healthy- yes, p = 0.049
total_volume_lost_dep_vs_healthy_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 27%, mean non_dep = 24%, p = 0.049

df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
names(df_for_visreg) <- c("EMPI", "prop_lost", "Diagnosis")
df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)

plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = prop_lost, col=Diagnosis)) +
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(0.05)) + 
  geom_boxplot(width=0.1) + theme_minimal() + 
  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed"))

print(plot)

#bar plot
Diagnosis <- c("MS w/Depression", "MS No Depression")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1))))

data <- data.frame(Diagnosis,value, sem)
#data$Fascicle <- pretty_names

# Grouped
barplot <- ggplot(data, aes(x=Diagnosis,fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", position=position_dodge(0.9)) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
  ylab("Proportion of Fascicles Affected") +
  ylim(0,0.5) +
  theme(axis.text = element_text(size = 12, colour = "black"))

print(barplot)

# Grouped
barplot <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Proportion of Fascicles Affected") +
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph

print(barplot)

#look at within depression network specifically - trend, p = 0.059
total_volume_lost_dep_vs_healthy_unique_in_dep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 36%, mean non_dep = 32%, p = 0.071

#look at within depression network in fascicles where at least 10% of the volume of the fascicle was in dep network
#yes, p = 0.048
total_volume_lost_dep_vs_healthy_unique_in_dep_network_10percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 33%, mean non_dep = 27%, p = 0.048

#look at within depression network in fascicles where at least 5% of the volume of the fascicle was in dep network
#yes, 
total_volume_lost_dep_vs_healthy_unique_in_dep_network_5percent <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 32%, mean non_dep = 27%, p = 0.056

#look at outside depression network, yes p = 0.037
total_volume_lost_dep_vs_healthy_unique_in_nondep_network <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2], var.equal = F) #mean dep = 18%, mean non_dep = 16%, p = 0.037

### Age actually makes things worse
#dep_vs_healthy_w_age <- lm(total_fascicle_prop_lost ~ depGroupVar + PAT_AGE_AT_EXAM, data = dep_and_healthy_groups_for_ICD_analysis)
#summary(dep_vs_healthy_w_age)

#df for visreg - actually not using visreg. Just a nice way to see that there is significantly more loss in depression network versus outside depression network

```

```{r depressed_versus_healthy_unique_subjs_indiv_fascicle}
##### repeat above analysis but use only the first instance of each EMPI, sorted by date
df_unique_empi <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#lm
fascicle_lm_unique <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM, list(i = as.name(x))), data = df_unique_empi)
})
names(fascicle_lm_unique) <- fascicle_names

#anova
fascicle_anova_unique <- lapply(fascicle_lm_unique, anova)

#fdr corrected
fascicle_anova_unique_fdr <- fdr_anova_generic(fascicle_anova_unique, 1)
print(fascicle_anova_unique_fdr)#anova values < 0.05, uncorrected

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr)

#fdr corrected only dep 10% mask
fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_bundle_mapping$inDepNetwork_vector ==2)]
fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr)

#fdr corrected only dep 5% mask
fascicle_anova_w_fiber_mapping <- fascicle_anova_unique[which(fascicle_volumes_dep_network$prop_in_mask >= 0.05)]
fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr)

#uncorrected
fascicle_anova_unique_p <- sapply(fascicle_anova_unique, function(v) v$"Pr(>F)"[1])
fascicle_anova_unique_p05_unc <- as.data.frame(fascicle_anova_unique_p[fascicle_anova_unique_p < 0.05])
print(fascicle_anova_unique_p05_unc)

### plot AF_R

df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
names(df_for_visreg) <- c("EMPI", "prop_lost", "Diagnosis")
df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)

plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = prop_lost, col=Diagnosis)) +
  ggtitle("Arcuate Fasciculus R") + 
 # geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(0.05)) + 
  geom_boxplot(width=0.1) + theme_minimal() + 
  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed"))

print(plot)

### Plot OR_R	0.69024050
df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$OR_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1], dep_and_healthy_groups_for_ICD_analysis$OR_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]), c(rep(1,298), rep(2,298))))
names(df_for_visreg) <- c("EMPI", "prop_lost", "Diagnosis")
df_for_visreg$Diagnosis <- as.factor(df_for_visreg$Diagnosis)

plot <- ggplot(df_for_visreg, aes(x=Diagnosis, y = prop_lost, col=Diagnosis)) +
  ggtitle("Optic Radiation R") + 
 # geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(0.05)) + 
  geom_boxplot(width=0.1) + theme_minimal() + 
  stat_summary(fun.y=median, geom="point", size=2, color="black") + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) 
print(plot)

### grouped bar plot

Diagnosis <- c(rep("Depressed" , 2) , rep("Non-Depressed" , 2))
Fascicle <- rep(c("Depression Network (AF)" , "Region Outside Depression Network (C_PH)"), 2)
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$C_PH_L[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2]),
           mean(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]),
           mean(dep_and_healthy_groups_for_ICD_analysis$C_PH_L[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$C_PH_L[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==2))),
           (sd(dep_and_healthy_groups_for_ICD_analysis$AF_R[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1 ))),
          (sd(dep_and_healthy_groups_for_ICD_analysis$C_PH_L[dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1])/sqrt(sum(dep_and_healthy_groups_for_ICD_analysis$depGroupVar ==1 ))))

data <- data.frame(Diagnosis,Fascicle,value, sem)
#data$Fascicle <- pretty_names

# Grouped
barplot <- ggplot(data, aes(fill=Diagnosis, y=value, x=Fascicle)) + 
    geom_bar(position="dodge", stat="identity") + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
  ylab("Proportion of Fascicles Affected") +
  theme(axis.text = element_text(size = 12, colour = "black"), axis.title.x = element_blank())

print(barplot)

# Grouped
barplot <- ggplot(data, aes(fill=Diagnosis, y=value, x=Fascicle)) + 
    geom_bar(position="dodge", stat="identity") + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
  ylab("Proportion of Fascicles Affected") +
  theme(axis.text = element_text(size = 12, colour = "black"), axis.title.x = element_blank())

print(barplot)

barplot2 <- ggplot(data, aes(fill=Diagnosis, y=value, x=Fascicle)) + 
    geom_bar(position="dodge", stat="identity", colour = "black", width=0.9,size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1) + 
    scale_fill_manual(values = c("#cdcdcd", "#FFFFFF", "#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph

    ylab("Proportion of Fascicles Affected") +
    theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
          axis.title.y = element_text(size=20),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty graph
    scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph

print(barplot2)

```

```{r testing_whether_ms_disease_more_likely_in_depression_mask_unique}
### tests whether there is higher proportion of vertices lost in depression network versus not

dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0)  %>% 
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
  mutate(total_fascicle_prop_lost_indepnet_10_percent =
           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
  mutate(total_fascicle_prop_lost_indepnet_5_percent =
           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
  mutate(total_fascicle_prop_lost_nondep_network =
           rowMeans(across(fascicle_names_nondep_network))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ### when I filter and remove fascicles where there was no loss at all, the depression vs healthy findings go away
#  filter(total_fascicle_prop_lost > 0) %>%
  ungroup()

#test whether there is differentially more volume loss in depression versus nondepression network
volume_lost_dep_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep_net = 35%, mean non_dep_net =  17%, p 2.2 x 10^-16

volume_lost_dep10_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_10_percent,
dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 31%, mean non_dep_net = 17%, p 2.2 * 10^-16

volume_lost_dep5_vs_nondep_network_unique <- t.test(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet_5_percent,
dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network, var.equal = F) #mean dep net = 30%, mean non_dep_net = 17%, p 2.2 * 10^-16

#df for visreg - actually not using visreg. Just a nice way to see that there is significantly more loss in depression network versus outside depression network
df_for_visreg <- as.data.frame(cbind(c(dep_and_healthy_groups_for_ICD_analysis$EMPI, dep_and_healthy_groups_for_ICD_analysis$EMPI), c(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet, dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network), c(rep(1,298), rep(2,298))))
names(df_for_visreg) <- c("EMPI", "prop_lost", "network")
df_for_visreg$network <- as.factor(df_for_visreg$network)

plot <- ggplot(df_for_visreg, aes(x=network, y = prop_lost, col=network)) +
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(0.05)) + 
  geom_boxplot(width=0.1) + theme_minimal() + 
  stat_summary(fun.y=median, geom="point", size=2, color="red") + 
  scale_x_discrete(labels=c("1" = "Depression Network", "2" = "Non-Depression Network")) + 
  ylab("Proportion of Fascicles Affected") + 
  theme(axis.text =element_text(size=14), axis.title.x = element_blank())

print(plot)

########## Bar plot
Network <- c("Depression Network","Non-Depression Network")
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet),
           mean(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_indepnet)/sqrt(dim(dep_and_healthy_groups_for_ICD_analysis)[1])),
           (sd(dep_and_healthy_groups_for_ICD_analysis$total_fascicle_prop_lost_nondep_network)/sqrt(dim(dep_and_healthy_groups_for_ICD_analysis)[1])))

data <- data.frame(Network,value,sem)
#data$Fascicle <- pretty_names

# Grouped
barplot <- ggplot(data, aes(x=Network,fill=Network, y=value)) + 
  geom_bar(stat="identity", position=position_dodge(0.9)) + 
  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
  ylab("Proportion of Fascicles Affected") +
  ylim(0,0.5) +
  theme(axis.text = element_text(size = 12, colour = "black"))

print(barplot)
# Black and white
barplot <- ggplot(data, aes(x=Network,fill=Network, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
  geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ylab("Proportion of Fascicles Affected") +
  theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        panel.background = element_rect(colour = "white", fill = "white")) + #comment out if you want pretty 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph

print(barplot)

###### Depression network x depression group interaction - It is there, depression always worse. 
#these lms largely match up with the t.tests above. Sig differences between depressed (more disease) within and outside of mask

loss_in_dep_network <- lm(total_fascicle_prop_lost_indepnet~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #non sig, p = 0.061
loss_in_dep_network_10_percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p = 0.0492
loss_in_nondep_network <- lm(total_fascicle_prop_lost_nondep_network~depGroupVar, data = dep_and_healthy_groups_for_ICD_analysis) #p=0.037
```

```{r phq9_correlation_fascicles}

####### NOTHING SIGNIFICANT!!!! ########3
#isolate out only the depressed group and healthy group
with_phq9 <- df_demo_and_fascicles %>%
  filter(depGroupVar != 0) %>%
  filter(!is.na(PHQ.9)) %>%
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
  mutate(total_fascicle_prop_lost_indepnet_10_percent =
           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
  mutate(total_fascicle_prop_lost_indepnet_5_percent =
           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
  mutate(total_fascicle_prop_lost_nondep_network =
           rowMeans(across(fascicle_names_nondep_network))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#lm
fascicle_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ PHQ.9, list(i = as.name(x))), data = with_phq9)
})
names(fascicle_lm) <- fascicle_names

#anova
fascicle_anova <- lapply(fascicle_lm, anova)

#fdr corrected
fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
print(fascicle_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr)

#fdr corrected only dep 10% mask
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr)

#uncorrected
fascicle_anova_p <- sapply(fascicle_anova, function(v) v$"Pr(>F)"[1])
fascicle_anova_p05_unc <- fascicle_anova_p[!is.na(fascicle_anova_p) & fascicle_anova_p < 0.05]
print(fascicle_anova_p05_unc)

######### overall fascicle proportion affected - nothing is significant
overall_prop_affected <- lm(total_fascicle_prop_lost~PHQ.9, data=with_phq9)
overall_prop_affected_depnet <- lm(total_fascicle_prop_lost_indepnet~PHQ.9, data=with_phq9)
overall_prop_affected_depnet_10percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~PHQ.9, data=with_phq9)
overall_prop_affected_depnet_5percent <- lm(total_fascicle_prop_lost_indepnet_5_percent~PHQ.9, data=with_phq9)
overall_prop_affected_nondep <- lm(total_fascicle_prop_lost_nondep_network~PHQ.9, data=with_phq9)
```

```{r phq9_regressions_just_dep_group}
##### NOTHING SIGNIFICANT ########
##################################
## Just within depression group ##
##################################
#isolate out only the depressed group and healthy group
with_phq9 <- df_demo_and_fascicles %>%
  filter(depGroupVar == 2) %>%
  filter(!is.na(PHQ.9)) %>%
  mutate(total_fascicle_prop_lost = rowMeans(across(fascicle_names))) %>%
  mutate(total_fascicle_prop_lost_indepnet = rowMeans(across(fascicle_names_dep_network_all))) %>%
  mutate(total_fascicle_prop_lost_indepnet_10_percent =
           rowMeans(across(fascicle_names_dep_network_10_percent))) %>%
  mutate(total_fascicle_prop_lost_indepnet_5_percent =
           rowMeans(across(fascicle_names_dep_network_5_percent))) %>%
  mutate(total_fascicle_prop_lost_nondep_network =
           rowMeans(across(fascicle_names_nondep_network))) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()
#lm
fascicle_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ PHQ.9, list(i = as.name(x))), data = with_phq9)
})
names(fascicle_lm) <- fascicle_names

#anova
fascicle_anova <- lapply(fascicle_lm, anova)

#fdr corrected
fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova, 1)
print(fascicle_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr<- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr)

#fdr corrected only dep 10% mask
fascicle_anova_w_fiber_mapping <- fascicle_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr)

#uncorrected
fascicle_anova_p <- sapply(fascicle_anova, function(v) v$"Pr(>F)"[1])
fascicle_anova_p05_unc <- fascicle_anova_p[!is.na(fascicle_anova_p) & fascicle_anova_p < 0.05]
print(fascicle_anova_p05_unc)

######### overall fascicle proportion affected - nothing is significant
overall_prop_affected <- lm(total_fascicle_prop_lost~PHQ.9, data=with_phq9)
overall_prop_affected_depnet <- lm(total_fascicle_prop_lost_indepnet~PHQ.9, data=with_phq9)
overall_prop_affected_depnet_10percent <- lm(total_fascicle_prop_lost_indepnet_10_percent~PHQ.9, data=with_phq9)
overall_prop_affected_depnet_5percent <- lm(total_fascicle_prop_lost_indepnet_5_percent~PHQ.9, data=with_phq9)
overall_prop_affected_nondep <- lm(total_fascicle_prop_lost_nondep_network~PHQ.9, data=with_phq9)

#visreg
#sapply(fascicle_lm, visreg)
```
All the mixed models are below
```{r age_effects_indiv_fascicles_mm}

### This subscript compares proportion of fascicles lost across age for each fascicle of the 80+. Looks lonly at the people in the depressed versus healthy group. Importantly, I ran this looking at everyone, at the results are the same.
### Separately considers findings just within the association cortex

#isolate out only the depressed group and healthy group
df_mm <- df_demo_and_fascicles %>% filter(depGroupVar != 0)
#n depressed = 859, healthy = 604, total n = 1463

#lm
fascicle_lm_mm <- lapply(fascicle_names, function(x) 
{
  lmerTest::lmer(substitute(i ~ PAT_AGE_AT_EXAM + (1 | EMPI), list(i = as.name(x))), data = df_mm)
})
names(fascicle_lm_mm) <- fascicle_names

#anova
fascicle_anova_mm <- lapply(fascicle_lm_mm, anova)

#fdr corrected
fascicle_anova_mm_fdr <- fdr_anova_generic(fascicle_anova_mm, 1)
print(fascicle_anova_mm_fdr)#anova values < 0.05, uncorrected

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network 10%
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_anova_mm_p <- sapply(fascicle_anova_mm, function(v) v$"Pr(>F)"[1])
fascicle_anova_mm_p05_unc <- as.data.frame(fascicle_anova_mm_p[fascicle_anova_mm_p < 0.05])
print(fascicle_anova_mm_p05_unc)

```

```{r sex_effects_fascicles_mm}
########### sex effect whole group

#isolate out only the depressed group and healthy group
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% filter(depGroupVar != 0) 
#healthhy = 713, total n = 1106

#### sex histo ###
hist(dep_and_healthy_groups_for_ICD_analysis$sex_binarized,
main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1), "/", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)),  xlab = "Sex", breaks = 2)

#lm
fascicle_sex_lm <- lapply(fascicle_names, function(x) 
{
  lmerTest::lmer(substitute(i ~ osex + (1|EMPI), list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_sex_lm) <- fascicle_names

#anova
fascicle_sex_anova <- lapply(fascicle_sex_lm, anova)

#fdr corrected
fascicle_sex_anova_fdr <- fdr_anova_generic(fascicle_sex_anova, 1)
print(fascicle_sex_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sex_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#visreg
#sapply(fascicle_sex_lm,visreg)
```

```{r depressed_versus_healthy_indiv_fascicles_mm}

### This subscript compares proportion of fascicles lost for each fascicle of the 80+ between those in the depressed and the clean healthy group. Done as both a mixed model as well as using unique individuals alone.

#Of note, adding age into this model makes things WORSE

### Separately considers findings just within the association cortex
#isolate out only the depressed group and healthy group
df_mm <- df_demo_and_fascicles %>% filter(depGroupVar != 0)
#n depressed = 859, healthy = 604, total n = 1463

#lm
fascicle_lm_mm <- lapply(fascicle_names, function(x) 
{
  lmerTest::lmer(substitute(i ~ depGroupVar + (1 | EMPI), list(i = as.name(x))), data = df_mm)
})
names(fascicle_lm_mm) <- fascicle_names

#anova
fascicle_anova_mm <- lapply(fascicle_lm_mm, anova)

#fdr corrected
fascicle_anova_mm_fdr <- fdr_anova_generic(fascicle_anova_mm, 1)
print(fascicle_anova_mm_fdr)#anova values < 0.05, uncorrected

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network 10%
fascicle_anova_w_fiber_mapping <- fascicle_anova_mm[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_anova_mm_p <- sapply(fascicle_anova_mm, function(v) v$"Pr(>F)"[1])
fascicle_anova_mm_p05_unc <- as.data.frame(fascicle_anova_mm_p[fascicle_anova_mm_p < 0.05])
print(fascicle_anova_mm_p05_unc)

```

Lastly, this section will explore interactions (age * dep), (sex * dep)

```{r age_by_depression_unique}

#isolate out only the depressed group and healthy group
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>%
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#lm
fascicle_agedep_lm <- lapply(fascicle_names, function(x) 
{
   lm(substitute(i ~ PAT_AGE_AT_EXAM*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
# lm(substitute(i ~ depGroupVar + PAT_AGE_AT_EXAM + osex, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_agedep_lm) <- fascicle_names

#anova
fascicle_agedep_anova <- lapply(fascicle_agedep_lm, anova)

#fdr corrected - depression alone
fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 2)
print(fascicle_agedep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#fdr corrected - interaction
fascicle_agedep_anova_fdr <- fdr_anova_generic(fascicle_agedep_anova, 3)
print(fascicle_agedep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network 10%
fascicle_anova_w_fiber_mapping <- fascicle_agedep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_anova_p <- sapply(fascicle_agedep_anova, function(v) v$"Pr(>F)"[3])
fascicle_anova_p05_unc <- as.data.frame(fascicle_anova_p[fascicle_anova_p < 0.05])
print(fascicle_anova_p05_unc)
```

```{r sex_by_depression_unique}

#isolate out only the depressed group and healthy group
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>%
  filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#lm
fascicle_sexdep_lm <- lapply(fascicle_names, function(x) 
{
   lm(substitute(i ~ osex*depGroupVar, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)

})
names(fascicle_sexdep_lm) <- fascicle_names

#anova
fascicle_sexdep_anova <- lapply(fascicle_sexdep_lm, anova)

#fdr corrected - depression alone
fascicle_sexdep_anova_fdr <- fdr_anova_generic(fascicle_sexdep_anova, 2)
print(fascicle_sexdep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_anova_p <- sapply(fascicle_sexdep_anova, function(v) v$"Pr(>F)"[2])
fascicle_anova_p05_unc <- as.data.frame(fascicle_anova_p[fascicle_anova_p < 0.05])
print(fascicle_anova_p05_unc)

#fdr corrected - interaction
fascicle_sexdep_anova_fdr <- fdr_anova_generic(fascicle_sexdep_anova, 3)
print(fascicle_sexdep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network
fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)

#fdr corrected only dep network 10%
fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector == 2)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,3)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_anova_p <- sapply(fascicle_sexdep_anova, function(v) v$"Pr(>F)"[3])
fascicle_anova_p05_unc <- as.data.frame(fascicle_anova_p[fascicle_anova_p < 0.05])
print(fascicle_anova_p05_unc)
```
Interaction models below

```{r sexdep_effects_fascicles_unique}

### Trying to figure out if there are any sex * depression effects, covarying for age given that men tend to get worse over time. dep*sex uncorrected in SLF1_L and CNVIII_R p< 0.05)

#isolate out only the depressed group and healthy group
dep_and_healthy_groups_for_ICD_analysis <- df_demo_and_fascicles %>% filter(depGroupVar != 0)  %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup()

#### sex histo ###
hist(dep_and_healthy_groups_for_ICD_analysis$sex_binarized,
main = paste0("Sex In Good Mimosa Group : n (M/F)= ", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1), "/", sum(dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)),  xlab = "Sex", breaks = 2)

#lm
fascicle_sexdep_lm <- lapply(fascicle_names, function(x) 
{
  lm(substitute(i ~ depGroupVar*osex + PAT_AGE_AT_EXAM, list(i = as.name(x))), data = dep_and_healthy_groups_for_ICD_analysis)
})
names(fascicle_sexdep_lm) <- fascicle_names

#anova
fascicle_sexdep_anova <- lapply(fascicle_sexdep_lm, anova)

##### depGroupVar ####
### 12 areas, nearly identical to reg model

#fdr corrected
fascicle_sexdep_anova_fdr <- fdr_anova_generic(fascicle_sexdep_anova, 1)
print(fascicle_sexdep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,1)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_sexdep_anova_p <- sapply(fascicle_sexdep_anova, function(v) v$"Pr(>F)"[1])
fascicle_sexdep_anova_p05_unc <- as.data.frame(fascicle_sexdep_anova_p[fascicle_sexdep_anova_p < 0.05])
print(fascicle_sexdep_anova_p05_unc) ### CNVIII p < 0.05

##### sex differences ####
## same as above, just add in CNVIII_R

#fdr corrected
fascicle_sexdep_anova_fdr <- fdr_anova_generic(fascicle_sexdep_anova, 2)
print(fascicle_sexdep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,2)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_sexdep_anova_p <- sapply(fascicle_sexdep_anova, function(v) v$"Pr(>F)"[2])
fascicle_sexdep_anova_p05_unc <- as.data.frame(fascicle_sexdep_anova_p[fascicle_sexdep_anova_p < 0.05])
print(fascicle_sexdep_anova_p05_unc) ### CNVIII p < 0.05

#######interaction ####

### SLF1_L and CNVIII_R

#fdr corrected interaction
fascicle_sexdep_anova_fdr <- fdr_anova_generic(fascicle_sexdep_anova, 4)
print(fascicle_sexdep_anova_fdr)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$name_vector == "association")]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,4)
print(fascicle_anova_fdr_association)

#fdr corrected only association cortex
fascicle_anova_w_fiber_mapping <- fascicle_sexdep_anova[which(fascicle_bundle_mapping$inDepNetwork_vector >0)]
fascicle_anova_fdr_association <- fdr_anova_generic(fascicle_anova_w_fiber_mapping,4)
print(fascicle_anova_fdr_association)

#uncorrected
fascicle_sexdep_anova_p <- sapply(fascicle_sexdep_anova, function(v) v$"Pr(>F)"[4])
fascicle_sexdep_anova_p05_unc <- as.data.frame(fascicle_sexdep_anova_p[fascicle_sexdep_anova_p < 0.05])
print(fascicle_sexdep_anova_p05_unc) #p uncorrected slf1_L p = 0.0488, CNVIII_R 0.01)

### grouped bar and line plot

Diagnosis <- c(rep("Depressed" , 2) , rep("Non-Depressed" , 2))
Sex <- rep(c("Male" , "Female"), 2)
value <- c(mean(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)]),
           mean(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)]),
           mean(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)]),
           mean(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)]))

sem <- c((sd(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)])/sqrt(sum((dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)))),
         (sd(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)])/sqrt(sum((dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 2) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)))),
         (sd(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)])/sqrt(sum((dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 1)))),
         (sd(dep_and_healthy_groups_for_ICD_analysis$SLF1_L[(dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)])/sqrt(sum((dep_and_healthy_groups_for_ICD_analysis$depGroupVar == 1) & (dep_and_healthy_groups_for_ICD_analysis$sex_binarized == 2)))))

data <- data.frame(Diagnosis,Sex,value, sem)
#data$Fascicle <- pretty_names

# Grouped
barplot <- ggplot(data, aes(fill=Sex, y=value, x=Diagnosis)) + 
    geom_bar(position="dodge", stat="identity") + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9)) + 
  ylab("Proportion of SLF Affected") +
  theme(axis.text = element_text(size = 12, colour = "black"), axis.title.x = element_blank())

print(barplot)

lineplot <- ggplot(data=data, aes(fill=Sex, y=value, x=Diagnosis, group = Sex)) +

    geom_line(aes(color=Sex, size=0.9)) +
    geom_point(aes(color=Sex, size=0.9)) +
    ylim(0.05,0.30)+
    ylab("Proportion of SLF Affected") +
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1) + 
    theme(axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank())

print(lineplot)

```