### 2024-08-30 scripts/FINAL_scripts_clean/MSAnxiety_manuscript_post_replication_prop_FINAL.Rmd
added. new plots

---
title: "ACNP 2024 MS ANXIETY"
author: "Erica Baller"
date: "7/1/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(stringr)
library(MatchIt)
library(ggpubr)
library(tidymv)
library(viridis)  # Ensure you have the viridis package
#functions
source("~/BBL/msanxiety/scripts/ms_functions.R")
```

## MS anxiety analyses

## R Markdown

The first chunk here is reading in appropriate data frames and preprocessing. Two big data frames are read in. The first is the demographic, drug, lab info for the MS patients (from the DAC pull). The second is the data frame of fascicle proportion affected for people with good mimosa (# volume affected/total volume of fascicle). The most important preprocessing occurs in a series of mutate commands. These essentially add new columns to the data frame that can later be used. The thrust of these mutate commands is to define who has depression and who does not.

Of note, in prior iterations of this analysis, we did not consider psychiatric medications in determining group inclusion/exclusion. Turns out it makes a difference, as a lot of people on psychotropic medications were counted as "controls/healthy" in previous analyses. For these preliminary analysis, our inclusion/exclusion are more stringent. Of note, the graph below is irrespective of whether they had good/bad MIMoSA. As such, the MIMoSA subset is smaller. It does however highlight the

<img width="70%" src="/Users/eballer/BBL/msdepression/data/drugs_data/MS_Patient_Depression_Categorizations.png"/>

Inclusion criteria below:

Inclusion MS Criteria (dataframe read in here has already been filtered down using the exclusions below previously)
  1) Scanned under MS protocol (42K)
  2) MS Diagnosis (17K)
  3) Seen by MS doctor (16K)

Inclusion criteria for Anxious Group
  1) Have an ICD10 F4* code (in our sample, people have F41.1; 41.3; 41.8; 41.9)
        (may use this later)- https://www.nami.org/About-Mental-Illness/Treatments/Mental-Health-Medications

Inclusion criteria for Healthy Group
  1) No F* diagnoses
  2) At least 1 PHQ2 or PHQ9 with a documented 0
  3) Free of psychiatric medications per NAMI website (plus editions per Dr. Baller)

Exclusions
  If you received the MS protocol or had a diagnosis of MS but never actually saw an MS provider, we cannot be sure that the diagnosis of MS is truly valid, as it is hard to diagnose anyway.
  If you did not have a diagnosis of depression and never completed a PHQ2 or 9, and had no history of medications, we could not safely confirm nor deny the presence of anxiety. These subjects were excluded from future analyses. We could consider adding these to the healthy group maybe, but not too inclined to do that right now.

Imaging exclusion
  We are currently only using people who had good MIMOSA (75 or 100). That severely limits our numbers, but that is okay! When we re-QC, we will probably get a lot more into this group.

Groups:
1. Total unique individuals who also have MS - 1026
2. Total unique who had good MIMoSA - 867 (lost 159)
3. Remove people without MS provider, n=814 (lost 53)

#figure brainstorming
1. Flowchart of inclusion/exclusion (825 participants -> 11 lost for poor streamline filtering -> 814 go into phenotyping) -> Elena
   A. Anxiety diagnosis
    I. Yes
       - Anxiety Med
          a. Yes -> MS+severeA (24 individuals)
          b. No  -> MS+mildA (87 indiv)
     II. No
        - Anxiety Med
          a. Yes -> MS+mildA (162)
          b. No
            - Has PHQ2 or 9 and 0 
              i. Yes-> MS+noA (99 individuals)
              ii. No -> Unclassified (442 individuals)

Final groups (MS+noA = 99, MS+mildA = 249, MS+severeA = 24)

2. Demographics (age, sex, race, +/- PHQ2, PHQ9, has anxiety dx, on anxiety meds, both, on dmt) -> elena to make all the csvs into tables
3. PROMIS scores (each type of measure is a line, y axis is score, x axis is anxiety dose)
4. Some sort of schematic of pipeline (Maggie) - 
  a. amygdala - mpfc blobs
  b. uncinate connecting the two
  c. white matter lesion dropped in uncinates
  d. streamline filtering
  e. mean of both side
5. Mean UF Anxiety diagnosis either violin or bar
6. Mean Anxiety dose either violin or bar

7. Supplement: 
 - More detailed streamline filtering
 - Depression group uncinate burden
 - Fornix diagnosis and dose; depression group 
 - Corr mean_UF with social problems?

## Make final data frame for analysis
### read in emr data
```{r read_in_df}
homedir <- "/Users/eballer/BBL/msanxiety/"

#open Rds, n=51238
data_2023_pull <- readRDS(paste0(homedir, "/data/data_2023_pull_with_curated_icd10_codes.rds"))
```

### read in Mimosa data
```{r read_in_mimosa_qc}
#This section checks how many unique individuals had QA data, and then how many met MIMoSA QC. Some additional people lost at registration.

#read in mimosa data
mimosa_qc <- read.csv("/Users/eballer/BBL/msdepression/data/melissa_martin_files/csv/mimosa_dataframe", sep = ",", header = FALSE) #n=2841

#preprocess data frame to extract dates, empis, and score, n=2841
empis_with_qc <- mimosa_qc %>% 
  mutate(EMPI = as.factor(gsub(x = V5, pattern = ".*sub-(.*)/ses-.*/.*", replacement="\\1", perl = TRUE))) %>%
  mutate(EXAM_DATE = gsub(x=V5, pattern = ".*/ses-(.{8})/.*", replacement = "\\1\\2\\3")) %>% 
  mutate(score = V3)

#merge with whole data frame to make sure we exclude people w/out MS
mimosa_qc_df <- merge(data_2023_pull, empis_with_qc, by = c("EMPI", "EXAM_DATE")) #n = 11252

#get list of unique patients, n = 1026
mimosa_qc_unique <- mimosa_qc_df %>%
  group_by(EMPI) %>%
  arrange(EXAM_DATE) %>%
  slice(1) %>%
  ungroup

#get list of unique patients with good QC, n = 867
mimosa_unique_score_gre_75 <- mimosa_qc_unique %>%
  filter(score >= 75) %>%
  group_by(EMPI) %>%
  arrange(EXAM_DATE) %>%
  mutate(EXAM_DATA = as.character(EXAM_DATE)) %>%
  mutate(EMPI = as.character(EMPI)) %>%
  dplyr::select(EMPI, EXAM_DATE, score) %>%
  slice(1) %>%
  ungroup 
```

### merge with Mimosa data
```{r merge_df_with_mimosa}

#n=867, individual scans
df_people_with_good_mimosa <- data_2023_pull %>% 
  group_by(EMPI, EXAM_DATE) %>%
  slice(1) %>%
  left_join(mimosa_unique_score_gre_75, by = c("EMPI", "EXAM_DATE")) %>%
  filter(!is.na(score))
```

### read in fascicle info, use this to determine proportion and volume of injury within each fascicle, and save it in the df_demo_and_fascicles_df
```{r read_in_fascicle_info}

### by the end of this chunk, the df should include 77 columns each for proportion of injury and volume of injury in each fascicle

##########################################################
#             Read in Fascicle info             #
##########################################################

fascicle_proportions <- read.csv(paste0("/Users/eballer/BBL/msdepression//results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
fascicle_proportions <- fascicle_proportions %>%
  mutate(EMPI = as.character(EMPI)) %>%
  mutate(EXAM_DATE = as.character(EXAM_DATE))

```

### merge data with good mimosa
```{r merge_with_good_mimosa}
##########################################################
#             Merge with data_empi              #
##########################################################

#n = 814
df_demo_and_fascicles <- df_people_with_good_mimosa %>% 
  left_join(fascicle_proportions, by = c("EMPI", "EXAM_DATE")) %>%
    filter(!is.na(Has.MS.Provider)) %>% #remove people who don't have MS provider, n=825 (lost 52) 
    filter(!is.na(CNII_L))# %>% #this is a sample fascicle that should have a value. if it does not, it means something is usually wrong with whole person. Remove rows with NAs, n=814, lost 11

```

### read in lesion volume size
```{r read_in_lesion_volumes}
##########################################################
#             Read in Lesion Volumes                     #
##########################################################

# contains the volume of lesioned voxels, i.e. a sum of all the areas of each individual's binary mask that are 1

lesion_volumes <- read.csv("/Users/eballer/BBL/msdepression/results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes", sep = ",", header = TRUE) #n=2336, 958 unique, identical to fascicle unique
lesion_volumes <- lesion_volumes %>%
  mutate(EMPI = as.character(EMPI)) %>%
  mutate(EXAM_DATE = as.character(EXAM_DATE))

```

### merge with data frame with good mimosas
```{r merge_lesion_volume_w_mimosa_df}
##########################################################
#             Merge lesion volumes with mimosa df        #
##########################################################
#merge whole data set with fascicle proportions, n = 814
df_demo_and_fascicles <- df_demo_and_fascicles %>%
  left_join(lesion_volumes, by = c("EMPI", "EXAM_DATE"))

```

### make separate columns for proportional injury (prop), total volume injury (vol)
```{r add_suffix_to_each_measure}

########################################
## Read in Fiber Volume Values        ##
########################################

#read in fiber volume values
fiber_volume_values <- read.csv("/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv", sep=",", header = T)

#########################################
## Get vectors with names of fascicles ##
#########################################

#fascicle names are contained in all but the first 2 columns of the fascicle_proportions df
fascicle_names <- names(fascicle_proportions[3:dim(fascicle_proportions)[2]])
fascicle_names_no_cranial_nerves <- fascicle_names[grep("CN", fascicle_names, invert = TRUE)] #n=77, drop cranial nerves

fascicle_names_w_prop_suffix <- gsub(pattern = "(.*)", replacement = "\\1_prop", perl=TRUE, x = fascicle_names_no_cranial_nerves)
fascicle_names_w_vol_suffix <- gsub(pattern = "(.*)", replacement = "\\1_vol", perl=TRUE, x = fascicle_names_no_cranial_nerves)

#write.table(fascicle_names,"/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fascicle_names.csv", row.names = F, quote = F, col.names = F)

##########################################################
#  Add columns for volume and change proportion names    #
##########################################################

#remove cranial nerves
fiber_volume_values_no_cranial_nerves <- fiber_volume_values %>% filter(fascicle %in% fascicle_names_no_cranial_nerves) %>%
  slice(order(factor(fascicle, levels = fascicle_names_no_cranial_nerves)))

#extract data frame that only includes the fascicle proportion lost
df_fascicles_prop <- df_demo_and_fascicles[ , (names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)] 
df_everything_else <- df_demo_and_fascicles[ , !(names(df_demo_and_fascicles) %in% fascicle_names_no_cranial_nerves)]

#multiply each row by the vector of total volumes to extract volume of disease within fascicle
df_fascicles_volumes <- sweep(x = df_fascicles_prop, MARGIN = 2,                        # Sweep with Complex STATS
                  STATS = fiber_volume_values_no_cranial_nerves$volume_full, FUN = "*") %>%
  mutate_if(is.numeric, round) #round all values

# change name of df_fascicle_volumes to have the suffix, and add a suffix for the proportion measures
names(df_fascicles_prop) <- fascicle_names_w_prop_suffix
names(df_fascicles_volumes) <- fascicle_names_w_vol_suffix

#combine back with data_fram
#df_demo_and_fascicle_volumes_full <- cbind(df_everything_else, df_fascicle_volumes)
df_demo_and_fascicles <- cbind(df_everything_else, df_fascicles_prop, df_fascicles_volumes) #n=814
```

##Anxiety Network from neurosynth (association, not uniformity)
###read in file with fasciclexanxiety overlap, and define anxiety network
```{r read_in_neurosynth_anxiety_network}
neurosynth_anxiety_network_fascicle_overlap <- read.csv("/Users/eballer/BBL/msanxiety/results/streamline_volume_within_anxiety_network_association_neuro_maps_FDR_01.csv", header = T, sep = ",") %>% 
  filter(fascicle %in% fascicle_names_no_cranial_nerves) %>%
  slice(order(factor(fascicle, levels = fascicle_names_no_cranial_nerves))) %>% #get rid of cranial nerves up front 
  replace(is.na(.), 0)

# Total volume of healthy fascicles within brain
total_fascicle_volume <- sum(neurosynth_anxiety_network_fascicle_overlap$num_voxels_total)

```

###Add columns in the main data frame indicating overall injury in the whole network, in the top 25 network (vol and prop), and bottom 75%
```{r add_anxiety_and_dep_network_details_to_data_frame}

#all of this has been checked - making sure prop * vol network = sum of damage

df_demo_and_fascicles <- df_demo_and_fascicles %>%

  mutate(mean_fascicle_prop_lost_no_cranial_nerves = rowMeans(across(all_of(fascicle_names_w_prop_suffix)))) %>%
  mutate(mean_fascicle_volume_lost_no_cranial_nerves = rowMeans(across(all_of(fascicle_names_w_vol_suffix)))) %>%

  #sum of total volume lost and proportion of volume lost 
  mutate(sum_fascicle_vol_lost = rowSums(across(all_of(fascicle_names_w_vol_suffix)))) %>%
  mutate(proportion_volume_lost_per_total_network_size_sum = sum_fascicle_vol_lost/total_fascicle_volume) %>%

  ##mean_uncinate
  mutate(mean_UF_vol = (UF_L_vol + UF_R_vol)/2) %>%
  mutate(mean_UF_prop = (UF_L_prop + UF_R_prop)/2) %>%

  ##mean_fornix
  mutate(mean_fornix_volume = (F_L_vol + F_R_vol)/2) %>%
  mutate(mean_fornix_prop = (F_L_prop + F_R_prop)/2) %>%

  ungroup()

#neurosynth_anxiety_network_fascicle_overlap
```

### add PROMIS
```{r promis_data_handling}
##########################################################
#            Read in PROMIS Data              #
##########################################################
promis_data <- readRDS(paste0(homedir, "/data/dac/promis_data_20220824_fixed_physical_health_cleaned_n1813_unique_indiv_n532_2015-2022.rds"))
promis_score_names <- names(promis_data[7:16])

df_people_with_promis_data <- df_demo_and_fascicles %>%
  filter(EMPI %in% promis_data$EMPI) #n=129

unique_empis_in_promis_group <- unique(df_people_with_promis_data$EMPI) #n=129

#go through each person w/promis data and append the columns to their row that reflects the closest collected PROMIS

#function, take an empi and date, returns the row of data w/promis scores
get_closest_promis_to_ms_exam_date <- function(subj_empi, subj_exam_date) {
  df_closest_promis <- promis_data %>%
    filter(EMPI == subj_empi) %>%
    mutate(distance_from_scan = abs(RECORDED_TIME-subj_exam_date)) %>%
    filter(abs(RECORDED_TIME-subj_exam_date)==min(abs(RECORDED_TIME-subj_exam_date))) %>%
    slice(1)
  return(df_closest_promis)
}

#make dataframe to store promis info
promis_info_closest_proximity <- data.frame(matrix(NA, nrow=dim(df_people_with_promis_data)[1], ncol=(dim(promis_data)[2] + 1)))
names(promis_info_closest_proximity) <- c(names(promis_data), "distance_from_scan")

for (line in 1:dim(df_people_with_promis_data)[1]) {
  #pull out subject empi and exam date
  subj_empi <- df_people_with_promis_data$EMPI[line]
  subj_exam_date <- df_people_with_promis_data$BEGIN_EXAM_DTTM[line]

  #returns whole line, promis scores, id, empi, everything
  closest_promis_data <- get_closest_promis_to_ms_exam_date(subj_empi = subj_empi, subj_exam_date = subj_exam_date)

  #store as new line
  promis_info_closest_proximity[line,] = closest_promis_data
}

#drop unnecessary columns c("DE_PAT_ID", "HUP_MRN", "SEX", "RACE")
promis_info_closest_proximity_just_scores <- promis_info_closest_proximity %>% 
	select(-c(DE_PAT_ID, HUP_MRN, SEX, RACE)) # include EMPI

df_people_with_promis_data_scores_added <- merge(df_people_with_promis_data, 
	promis_info_closest_proximity_just_scores, by = "EMPI")

#summarize the distance from scans
summary(df_people_with_promis_data_scores_added$distance_from_scan/364) #mean 3.1 ys, median 2.8 years, range 0.39 y to 7.16

hist(df_people_with_promis_data_scores_added$distance_from_scan/364, 
     main = "Hisogram of years between Promis and Scan",
     xlab = "Years from Scan") #convert to years

```

## Demographics
### simple demographics
```{r demographics}

#this will display a demographics table comparing age, race, sex between healthy and depressed groups, using the first presentation to treatment
#demo simple, all groups
simple_demo_all_groups <- make_demographics_table_ms_anxiety_simple(data_frame = df_demo_and_fascicles[df_demo_and_fascicles$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified",])

#save 
simple_demo_all_groups_df <- print(simple_demo_all_groups, as.data.frame = TRUE)
write.csv(simple_demo_all_groups_df, file = paste0(homedir, "/results/simple_demo_all_groups_df.csv"), row.names = TRUE)

#demo simple, just MS+noA and MS+severeA
simple_demo_diagnosis <- make_demographics_table_ms_anxiety_simple(data_frame = df_demo_and_fascicles[((df_demo_and_fascicles$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") & (df_demo_and_fascicles$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds")),])

#save 
simple_demo_diagnosis_df <- print(simple_demo_diagnosis, as.data.frame = TRUE)
write.csv(simple_demo_diagnosis_df, file = paste0(homedir, "/results/simple_demo_diagnosis_df.csv"), row.names = TRUE)
```

### demographics with PROMIS
```{r demo_with_promis}
#just good mimosa w/promis
df_people_with_promis_data_scores_added_unique <- df_people_with_promis_data_scores_added %>%
  filter(anxiety_dose != -1)  %>%
  group_by(EMPI) %>% 
  mutate(num_na = sum(is.na(across(promis_score_names)))) %>%
  filter(num_na == 0) %>%
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
simple_demo_promis_all_groups <- make_demographics_table_ms_w_promis_anxiety(data_frame = df_people_with_promis_data_scores_added_unique)

#save 
simple_demo_promis_all_groups_df <- print(simple_demo_promis_all_groups, as.data.frame = TRUE)
write.csv(simple_demo_promis_all_groups_df, file = paste0(homedir, "/results/simple_demo_promis_all_groups_df.csv"), row.names = TRUE)

simple_demo_promis_diagnosis <- make_demographics_table_ms_w_promis_anxiety(data_frame = df_people_with_promis_data_scores_added_unique[(df_people_with_promis_data_scores_added_unique$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds"),])

#save 
simple_demo_promis_diagnosis_df <- print(simple_demo_promis_diagnosis, as.data.frame = TRUE)
write.csv(simple_demo_all_groups_df, file = paste0(homedir, "/results/simple_demo_promis_diagnosis_df.csv"), row.names = TRUE)

#------

#testing whether PROMIS sub sample is equally representative with respect to prevalence of depression/non-depression. I.e. do more people with depression get PROMIS scores? or are we equally drawing from both groups?

#first number is # with severeA, second is number without anxiety
num_total <- c(sum(df_demo_and_fascicles$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed == "Has.Anxiety.And.On.Anxiety.Meds.Inc.Dep"), sum(df_demo_and_fascicles$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed == "true_healthy"))

#first number is # with severeA, second is number without anxiety
num_w_promis <- c(sum(df_people_with_promis_data_scores_added_unique$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed == "Has.Anxiety.And.On.Anxiety.Meds.Inc.Dep"), sum(df_people_with_promis_data_scores_added_unique$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed == "true_healthy"))

df_2x2_promis_score_by_total_people_per_diagnosis <- data.frame(cbind(num_w_promis, num_total), row.names = c("MS+severeA", "MS+NoA"))

#X-squared = 1.9699, df = 1, p-value = 0.1605, not significant); 
chisq <- chisq.test(df_2x2_promis_score_by_total_people_per_diagnosis)
print(chisq)

#fishers exact:  p-value = 0.1313, 95 % CI 0.7169704 7.4524239, OR 2.388734 
fisher_test <- fisher.test(df_2x2_promis_score_by_total_people_per_diagnosis)
print(fisher_test)
```
#### Plot promis x anxiety dose
```{r plot_Promis_x_anxiety}
#set mood versus physical domains
mood_domains <- c("Quality_of_Life", 
                "Mental_Health_and_Mood",
                "Social_Activities_Satisfaction",
                "Carrying_Out_Social_Activities",
                "Emotional_Problems", 
                "PostOp.PROMIS.Mental.Score")

physical_domains <- c("Physical_Health", 
                      "Carrying_Out_Physical_Activities",
                      "Fatique_Average",
                      "PostOp.PROMIS.Physical.Score")

# Reshape the data from wide to long format
df_for_promis_plot <- df_people_with_promis_data_scores_added_unique %>%
  dplyr::select(EMPI,
                anxiety_dose, 
                Quality_of_Life, 
                Physical_Health, 
                Mental_Health_and_Mood,
                Social_Activities_Satisfaction,
                Carrying_Out_Social_Activities,
                Carrying_Out_Physical_Activities,
                Emotional_Problems, 
                Fatique_Average,
                PostOp.PROMIS.Physical.Score,
                PostOp.PROMIS.Mental.Score) %>%
  group_by(anxiety_dose) %>%

  # summarize by score
  summarise(across(c(Quality_of_Life, 
                     Physical_Health, 
                     Mental_Health_and_Mood,
                     Social_Activities_Satisfaction,
                     Carrying_Out_Social_Activities,
                     Carrying_Out_Physical_Activities,
                     Emotional_Problems, 
                     Fatique_Average,
                     PostOp.PROMIS.Physical.Score, #this has a much bigger range
                     PostOp.PROMIS.Mental.Score), #bigger range
                   list(mean = ~mean(. , na.rm = TRUE)))) %>% #                   ,     sd = ~sd(. , na.rm = TRUE))))

  # Reshape to long format for plotting
    pivot_longer(cols = -anxiety_dose, 
               names_to = "PROMIS_Score", 
               values_to = "Score")

custom_colors <- c(
  "Quality_of_Life" = "#66c2a5",  # Green
  "Physical_Health" = "#8da0cb",  # Purple
  "Mental_Health_and_Mood" = "#99d8c9",  # Green
  "Social_Activities_Satisfaction" = "#238b45",  # Green
  "Carrying_Out_Social_Activities" = "#41ae76",  # Green
  "Carrying_Out_Physical_Activities" = "#8856a7",  # Purple
  "Emotional_Problems" = "#006d2c",  # Dark Green
  "Fatique_Average" = "#4B0082",  # Light Purple
  "PostOp.PROMIS.Mental.Score" = "#9ACD32",  # Green
  "PostOp.PROMIS.Physical.Score" = "#810f7c"  # Dark Purple
)

custom_colors_simple <- c("#8856a7","#41ae76", "#006d2c", "#4B0082", "#99d8c9", "#8da0cb",  "#9ACD32", "#810f7c", "#66c2a5", "#238b45") #done in alphabetical order.

# Create the ggplot
promis_with_post_op <- ggplot(df_for_promis_plot, aes(x = anxiety_dose, y = Score, color = PROMIS_Score)) +
  geom_line(size = 1.2) +
  scale_x_continuous(breaks = c(0, 1, 2)) +
  scale_y_continuous(limits = c(0, 20)) +
  theme_minimal() +
  labs(x = "Anxiety Dose", y = "Mean PROMIS Score", 
       title = "PROMIS Scores by Anxiety Dose") +
#  scale_color_brewer(palette = "Set3") 
  scale_color_manual(values = custom_colors_simple)
promis_with_post_op

#remove postop scores

promis_without_postop<- ggplot(df_for_promis_plot, aes(x = anxiety_dose, y = Score, color = PROMIS_Score)) +
  geom_line(size = 1.2) +
  scale_x_continuous(breaks = c(0, 1, 2)) +
  scale_y_continuous(limits = c(0, 5)) +
  theme_minimal() +
  labs(x = "Anxiety Dose", y = "Mean PROMIS Score", 
       title = "PROMIS Scores by Anxiety Dose") +
  #scale_color_brewer(palette = "Set3") 
  scale_color_manual(values = custom_colors_simple)

promis_without_postop

```

#### Plot 
#####Promis x anxiety dose, group by mental vs physical health
```{r plot_Promis_x_anxiety}
#set mood versus physical domains
mood_domains <- c("Mental_Health_and_Mood",
                "Social_Activities_Satisfaction",
                "Carrying_Out_Social_Activities",
                "Emotional_Problems", 
                "Quality_of_Life",
                "PostOp.PROMIS.Mental.Score")

physical_domains <- c("Physical_Health", 
                      "Carrying_Out_Physical_Activities",
                      "Fatique_Average",
                      "PostOp.PROMIS.Physical.Score")

# Reshape the data from wide to long format
df_for_promis_plot_mental_vs_physical_healthy <- df_people_with_promis_data_scores_added_unique %>%
  group_by(EMPI) %>%
  mutate("Physical Health Burden" = sum(across(physical_domains))/length(physical_domains)) %>%
  mutate("Mental Health Burden" = sum(across(mood_domains))/length(mood_domains)) %>%
  dplyr::select(EMPI,
                anxiety_dose,
                "Physical Health Burden",
                "Mental Health Burden") %>%
  gather(health_variable, PROMIS_Score, -anxiety_dose, -EMPI) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  group_by(anxiety_dose, health_variable) %>%
  summarise(mean_promis_score = mean(PROMIS_Score , na.rm = TRUE),
            sd_promis_score = sd(PROMIS_Score, na.rm = TRUE),
            sem_promis_score = sd(PROMIS_Score, na.rm = TRUE) / sqrt(n())) %>%
  ungroup()

#custom_colors_simple <- c("#9ACD32", "#4B0082") #done in alphabetical order.

# Define the desired order of legend items
desired_order <- c("Physical Health Burden", "Mental Health Burden")

# Ensure the factor levels are set to match the desired order
df_for_promis_plot_mental_vs_physical_healthy$health_variable <- factor(
  df_for_promis_plot_mental_vs_physical_healthy$health_variable,
  levels = desired_order
)

# Define custom colors, ensuring the names match the levels of the factor
custom_colors_simple <- c("Physical Health Burden" = "#4B0082",  # Purple
                          "Mental Health Burden" = "#9ACD32")  # Green

# Create the ggplot
promis_with_post_op_mental_vs_physical <- ggplot(df_for_promis_plot_mental_vs_physical_healthy, aes(x = anxiety_dose, y = mean_promis_score, color = health_variable)) +
  geom_line(size = 1.2) +
  scale_x_continuous(breaks = c(0, 1, 2),  # Keep numeric values on x-axis
                     labels = c("MS+noA", "MS+mildA", "MS+severeA")) +  # Set custom labels for each numeric value
  scale_y_continuous(limits = c(0, 7)) +
  geom_errorbar(aes(ymin = mean_promis_score - sem_promis_score, ymax = mean_promis_score + sem_promis_score), width = 0.2, size = 0.8)  +
  theme(
        axis.text.y = element_text(size=12, colour = "black"),
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size=12, colour = "black"),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.95),  # Move legend to upper right of plot area (0.9, 0.9 corresponds to top-right)
        legend.background = element_rect(fill = "white"),  # Set legend background to white
        legend.key = element_rect(fill = "white", color = "white"),  # Set legend key background to white
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white")) +
  labs(y= "Mean PROMIS Score") +
  scale_color_manual(values = custom_colors_simple, breaks = desired_order)

promis_with_post_op_mental_vs_physical
#save pdf
pdf(file=paste0(homedir, "/results/Figures_for_msanxiety_paper/promis_with_post_op_mental_vs_physical.pdf"))
  promis_with_post_op_mental_vs_physical
dev.off()

```

##### Physical and Mental Health x anxiety dose intx -N.S.
```{r phys_and_mental_x_anxiety_dose}
###
df_for_promis_plot_mental_vs_physical_healthy_no_summarize <- df_people_with_promis_data_scores_added_unique %>%
  group_by(EMPI) %>%
  mutate(physical_health_score = sum(across(physical_domains))/length(physical_domains)) %>%
  mutate(mental_health_score = sum(across(mood_domains))/length(mood_domains)) %>%
  mutate(anxiety_dose = as.numeric(anxiety_dose)) %>%
  dplyr::select(EMPI,
                anxiety_dose,
                physical_health_score,
                mental_health_score) %>%
  gather(health_variable, PROMIS_Score, -anxiety_dose, -EMPI)

#anxiety_dose                                       -1.1181     0.2816 88.0000  -3.970 0.000146 ***, others NS
anxiety_dose_by_health_var_score <- lmer(PROMIS_Score ~anxiety_dose*health_variable + (1|EMPI), df_for_promis_plot_mental_vs_physical_healthy_no_summarize)
summary(anxiety_dose_by_health_var_score)

### do just with super sick and healthy
df_for_promis_plot_mental_vs_physical_healthy_no_summarize_just_sick_and_healthy <- df_people_with_promis_data_scores_added_unique %>%
  group_by(EMPI) %>%
  filter(anxiety_dose != 1) %>%
  mutate(physical_health_score = sum(across(physical_domains))/length(physical_domains)) %>%
  mutate(mental_health_score = sum(across(mood_domains))/length(mood_domains)) %>%
  mutate(anxiety_dose = as.numeric(anxiety_dose)) %>%
  dplyr::select(EMPI,
                anxiety_dose,
                physical_health_score,
                mental_health_score) %>%
  gather(health_variable, PROMIS_Score, -anxiety_dose, -EMPI)

#aanxiety_dose                                       -1.0417     0.2611 24.3607  -3.989  0.00053 ***
##health_variablephysical_health_score                0.1875     0.2731 17.0000   0.687  0.50155    
#anxiety_dose:health_variablephysical_health_score   0.4063     0.2249 17.0000   1.806  0.08864 .  
anxiety_dose_by_health_var_score_just_sick_and_healthy <- lmer(PROMIS_Score ~anxiety_dose*health_variable + (1|EMPI), df_for_promis_plot_mental_vs_physical_healthy_no_summarize_just_sick_and_healthy)
summary(anxiety_dose_by_health_var_score_just_sick_and_healthy)
```

#COLLAPSE anxiety dx and antianxiety meds with those plus depression, and anxiety dx OR on anxiety meds into 1 group -->> THIS IS THE MONEY

### Double check that people in each group are correct
```{r gut_check}

#want to make sure that people in healthy are NOT on anxiolytics or have dx, that people in the second group have one OR the other, and that people in the third group have BOTH

#make data frame
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
   filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
   filter(anxiety_dose != 3)

###true healthy
num_people_in_healthy_with_0_for_Has.Anxiety_and_On.Anxiolytics <- length(which(df_demo_and_fascicles_no_unclass_anxiety_dose$On.Anxiolytics ==0 & df_demo_and_fascicles_no_unclass_anxiety_dose$Has.anxietydx ==0))
print(paste0("num_people_in_healthy_with_0_for_Has.Anxiety_and_On.Anxiolytics-> ", num_people_in_healthy_with_0_for_Has.Anxiety_and_On.Anxiolytics, " and num in this group from groupVar -> ", sum(df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 0)))

### Anxiety dose = 1
num_people_in_Anxiety_OR_On.Anxiolytics <- length(which((df_demo_and_fascicles_no_unclass_anxiety_dose$On.Anxiolytics ==1 | df_demo_and_fascicles_no_unclass_anxiety_dose$Has.anxietydx ==1) &
(df_demo_and_fascicles_no_unclass_anxiety_dose$On.Anxiolytics != df_demo_and_fascicles_no_unclass_anxiety_dose$Has.anxietydx)))

print(paste0("num_people_in_Anxiety_OR_On.Anxiolytics-> ", num_people_in_Anxiety_OR_On.Anxiolytics, " and num in this group from groupVar -> ", sum(df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 1)))

#severe anxiety
num_people_in_Anxiety_AND_Anxiolytics <- length(which(df_demo_and_fascicles_no_unclass_anxiety_dose$On.Anxiolytics ==1 & df_demo_and_fascicles_no_unclass_anxiety_dose$Has.anxietydx ==1))
print(paste0("num_people_in_Anxiety_AND_Anxiolytics-> ", num_people_in_Anxiety_AND_Anxiolytics, " and num in this group from groupVar -> ", sum(df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 2)))

print(paste0("total n -> ",num_people_in_healthy_with_0_for_Has.Anxiety_and_On.Anxiolytics + num_people_in_Anxiety_OR_On.Anxiolytics + num_people_in_Anxiety_AND_Anxiolytics, " and row # in data frame -> ", dim(df_demo_and_fascicles_no_unclass_anxiety_dose)[1]))

```

### Overlap with depression group
```{r overlap_with_dep_group}
#### 53.4% overlap total

#read csv of empis from depression paper
depression_empis <- read.table(file="/Users/eballer/BBL/msanxiety/data/empis_of_people_in_BP_depression_paper.csv")
names(depression_empis) <- "EMPI"

#get final anxiety dataset and make column if the empi is in the prev list
df_disease_burden_unc_and_anx_net_collapsed_variables <- df_demo_and_fascicles %>%
  mutate(EMPI = as.numeric(EMPI)) %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  mutate(EMPI_in_dep_group = EMPI %in% depression_empis$EMPI)

#"Percent overlap 53.2%"
print(paste0("Percent overlap ", round(sum(df_disease_burden_unc_and_anx_net_collapsed_variables$EMPI_in_dep_group)/length(df_disease_burden_unc_and_anx_net_collapsed_variables$EMPI), 3) * 100, "%"))

 tibble_overall_counts <- df_disease_burden_unc_and_anx_net_collapsed_variables %>% 
  group_by(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  summarize(n(),
            num_in_dep_group = sum(EMPI_in_dep_group),
            percent_overlap = num_in_dep_group/n())
 tibble_overall_counts

print(paste0("Percent overlap with depression group -> ", round(sum(tibble_overall_counts$num_in_dep_group)/sum(tibble_overall_counts$`n()`),2)))
```

### Overall Disease by Anxiety Dose *acnp
```{r group_anxietydx_meds_and_dep_comorbidity}
#this collapses anxiety group + meds with those with anxiety group + meds + with/without depression

df_disease_burden_unc_and_anx_net_collapsed_variables <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>% 
  dplyr::select(proportion_volume_lost_per_total_network_size_sum, UF_L_prop, UF_R_prop, oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed, osex)

 tibble_overall_summaries <- df_disease_burden_unc_and_anx_net_collapsed_variables %>% 
  group_by(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  summarize(n(),
            mean_disease_burden = mean(proportion_volume_lost_per_total_network_size_sum),
            mean_UF_L = mean(UF_L_prop),
            mean_UF_R = mean(UF_R_prop))
tibble_overall_summaries

tibble_overall_summaries_by_sex <- df_disease_burden_unc_and_anx_net_collapsed_variables %>% 
  group_by(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed, osex) %>%
  summarize(n(),
            mean_disease_burden = mean(proportion_volume_lost_per_total_network_size_sum),
            mean_UF_L = mean(UF_L_prop),
            mean_UF_R = mean(UF_R_prop))
  tibble_overall_summaries_by_sex

  tibble_overall_summaries_females <- df_disease_burden_unc_and_anx_net_collapsed_variables %>% 
  filter(osex == "Female") %>%
  group_by(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  summarize(n(),
            mean_disease_burden = mean(proportion_volume_lost_per_total_network_size_sum),
            mean_UF_L = mean(UF_L_prop),
            mean_UF_R = mean(UF_R_prop))
  tibble_overall_summaries_females

    tibble_overall_summaries_males <- df_disease_burden_unc_and_anx_net_collapsed_variables %>% 
  filter(osex == "Male") %>%
  group_by(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  summarize(n(),
            mean_disease_burden = mean(proportion_volume_lost_per_total_network_size_sum),
            mean_UF_L = mean(UF_L_prop),
            mean_UF_R = mean(UF_R_prop))
  tibble_overall_summaries_males

  #mean disease

  df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>% filter (anxiety_dose != -1) #-1 is unclassified, 0 is no anxiety, 1 is diagnosis or on meds, 2 is diagnosis AND on meds, -> n = 371

  #anxiety_dose  0.04849    0.01890   2.566   0.0107 *  
  anxietyGrouping_by_sex_mean_disease_gam <- gam(proportion_volume_lost_per_total_network_size_sum ~ anxiety_dose + osex+ s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_by_sex_mean_disease_gam)

  #print summary stats
  gam_summary_stats(anxietyGrouping_by_sex_mean_disease_gam)
```

### Lesion volumes 
#### Diagnosis 
  - sick > healthy in vol of lesions
```{r lesion_volumes_dx}
 df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, volume_of_mimosa_lesions) %>%
  ungroup() #n=123

##### healthy vs sick

#gam lesion volumes; DiagnosisMS+Anxiety   8215.7     2702.7   3.040  0.00292 ** 
volume_of_mimosa_lesions_gam <- gam(volume_of_mimosa_lesions ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(volume_of_mimosa_lesions_gam)

 #print summary stats
 gam_summary_stats(volume_of_mimosa_lesions_gam)
```

#### Anxiety dose
  - anxiety dose correlates with disease burden (volume of lesions)
```{r lesion_volumes_anxiety_dose}
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter (anxiety_dose != -1) #n=372

 #bilat anxiety dose + sex + age: anxiety_dose     3866       1254   3.084   0.0022 **

  anxiety_dose_volume_of_mimosa_lesions_by_sex_gam <- gam(volume_of_mimosa_lesions ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxiety_dose_volume_of_mimosa_lesions_by_sex_gam)

   #print summary stats
 gam_summary_stats(anxiety_dose_volume_of_mimosa_lesions_by_sex_gam)

```

### Uncinate **ACNP**
#### Presence of Uncinate Lesions chisq - N.S.
```{r uncinate_presence_of_lesions}
### Basically, I'd like to answer the question - is it just a presence or absence of uncinate lesions, or is it the burden of uncinate lesions/
### To do this, I perform a chisq to see if there is a statistically significant difference in the number of people with uncinate lesions within each of the diagnostic groups
### I found that there are no statistically significant differences in the frequency of uncinate lesions between groups - meaning that it is not just the presence of lesions, but the impact

#define group and make column for people who do have uncinate lesions and those who don't
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) %>% #n=372
  mutate(has.uncinate.lesion = ifelse(mean_UF_prop >0, 1, 0)) #anywhere where there is *some* overlap counts as a 1 - they have an uncinate lesion, otherwise 0

#define number of people within each group that have uncinate lesions
uncinate_lesions_dose_0 <- df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(has.uncinate.lesion == 1 & anxiety_dose == 0) %>% count() #n=76
uncinate_lesions_dose_1 <- df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(has.uncinate.lesion == 1 & anxiety_dose == 1) %>% count() #n=190
uncinate_lesions_dose_2 <- df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(has.uncinate.lesion == 1 & anxiety_dose == 2) %>% count() #n=20

#define total number of people in each group
anxiety_total_dose_0 <- df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(anxiety_dose == 0) %>% count() #n=99
anxiety_total_dose_1 <- df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(anxiety_dose == 1) %>% count() #n=249
anxiety_total_dose_2 <- df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(anxiety_dose == 2) %>% count() #n=24

#make table
num_w_uncinate_lesions <- as.numeric(c(uncinate_lesions_dose_0, uncinate_lesions_dose_1, uncinate_lesions_dose_2))
num_total <-as.numeric(c(anxiety_total_dose_0,anxiety_total_dose_1, anxiety_total_dose_2))
df_2x3_presence_of_UF_lesions_by_total_people_per_dose <- data.frame(cbind(num_w_uncinate_lesions, num_total), row.names = c("MS+noAnxiety","MS+AnxietyorMed", "MS+AnxietyandMed"))

#X-squared (Pearson's) X-squared(2,372) = 0.077073, p = 0.9622
chisq_anxiety_dose <- chisq.test(x=df_2x3_presence_of_UF_lesions_by_total_people_per_dose)
chisq_anxiety_dose

#just healthy and super anxious/sick
num_w_UF_just_healthy_and_sick <- as.numeric(c(uncinate_lesions_dose_0, uncinate_lesions_dose_2))
num_total_just_healthy_and_sick <-as.numeric(c(anxiety_total_dose_0,anxiety_total_dose_2))
df_2x2_UF_lesions_by_total_people_per_diagnosis_num_total_just_healthy_and_sick <- data.frame(cbind(num_w_UF_just_healthy_and_sick, num_total_just_healthy_and_sick), row.names = c("MS+noAnxiety", "MS+AnxietyandMed"))

#X-squared(1,123) = 0.0052081, p = 0.9425
chisq_healthy_vs_sick <- chisq.test(x = df_2x2_UF_lesions_by_total_people_per_diagnosis_num_total_just_healthy_and_sick)
chisq_healthy_vs_sick

```

#### Uncinate by anxiety diagnosis (vol and prop)
```{r uncinate_by_dx}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

#### Fx = F gives better results than FX = T

### Results:  
#define group
  uncinate_by_dx_df <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  ungroup() #n=123

#bilateral UF ~ Diagnosis: DiagnosisMS+Anxiety   0.10624    0.05255   2.022   0.0455  * 
disease_mean_UF_prop_gam <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_prop_gam)

 #print summary stats
  gam_summary_stats(disease_mean_UF_prop_gam)
```

#### Uncinate by Anxiety Dose 
```{r uncinateByAnxietyDose}

#### Fx = F gives better results than FX = T
#results: Anxiety dose related to UF_R when covarying for linear effects of age and sex, just barely
# Additionally - taking the mean of both UF and UR, LM for bilat anxiety dose + sex + age: anxiety_dose most sig P= 0.0367 *  (fx=T)

#define group
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) #n=372

    #bilat anxiety dose + sex + age: anxiety_dose    0.04989    0.02397   2.082   0.0381 * 
 anxietyGrouping_mean_UF_by_sex_gam_prop <- gam(mean_UF_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam_prop)

  #print summary stats
  gam_summary_stats(anxietyGrouping_mean_UF_by_sex_gam_prop)

```

###### Laterality in uncinate (prop)
```{r laterality}
  #laterality 
  #overall - t = -7.9761, df = 668.34, p-value = 6.564e-15
  right_vs_left_uncinate <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_prop, df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_prop)
  print(right_vs_left_uncinate)

  #healthy - t = -4.2898, df = 173.04, p-value = 2.969e-05
   right_vs_left_uncinate_healthy <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 0], df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 0])
  print(right_vs_left_uncinate_healthy)

  #anxietydx or meds, t = -6.3724, df = 447.66, p-value = 4.629e-10
  right_vs_left_uncinate_anxiety_or_meds <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 1], df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 1])
  print(right_vs_left_uncinate_anxiety_or_meds)

  #anxious+meds,t = -2.19, df = 43.079, p-value = 0.03398
   right_vs_left_uncinate_anxiety_and_meds <- t.test(df_demo_and_fascicles_no_unclass_anxiety_dose$UF_L_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 2], df_demo_and_fascicles_no_unclass_anxiety_dose$UF_R_prop[df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose == 2])
  print(right_vs_left_uncinate_anxiety_and_meds)
```

#### Sensitivity
###### UF by depGroupVar (vol and prop, N.S)
```{r UF_by_depGroupVar}
# Specificity - want to make sure the UF findings are specific to anxiety, so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

df_dep <- df_demo_and_fascicles %>%
  filter(depGroupVar != 0) %>%
  mutate(depDiagnosis = as.factor(ifelse(depGroupVar ==2, "MS+Depression", "MS+noDepression")))

table(df_dep$depDiagnosis) #MS+Dep n = 207; MS+noDep n = 99

 #mean UF, depDiagnosisMS+noDepression   -0.05029    0.03028  -1.661   0.0978 . 
dep_mean_UF_gam_prop <- gam(mean_UF_prop ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
summary(dep_mean_UF_gam_prop)

  #print summary stats
 gam_summary_stats(dep_mean_UF_gam_prop)
```

###### Fornix by Anxiety Diagnosis and Dose (vol and prop, N.S)
```{r fornixByAnxietyDose}

#results: Anxiety dose NOT related to F_R when covarying for linear effects of age and sex, not in line with that one paper that found fornix differences
#define group

#anxiety diagnosis
 df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  ungroup() #n=123

 ##### healthy vs sick - prop
#Fornix gam Diagnosis; DiagnosisMS+Anxiety    0.03690    0.05405   0.683   0.4961    - NS
anxietyDiagnosis_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(anxietyDiagnosis_mean_F_by_sex_gam_prop)

#define group for dose
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1)

  #gam, Fornix anxiety_dose   0.02303    0.02401   0.959    0.338    - NS
  anxietyGrouping_mean_F_by_sex_gam_prop <- gam(mean_fornix_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_F_by_sex_gam_prop)

 #print summary stats
 gam_summary_stats(anxietyDiagnosis_mean_F_by_sex_gam_prop)
 gam_summary_stats(anxietyGrouping_mean_F_by_sex_gam_prop)
```

##### No steroids/IF -N.S. but low n's
```{r no_steroids_or_if}
#### Steroids and interferon are known to have psychiatric side effects
### we repeated our analyses but removed these groups
### Nothing sig but trend, probably because we lose N (2 people in MS+noA, 1 in MS+mildA, 2 in MS+severeA)
### Will not report these analyses, but just note number of people in the table who are on these things

### remove people on IF or steroids
df_demo_and_fascicles_no_steroids_or_ifn <- df_demo_and_fascicles %>%
  filter((On.Steroids == 0) & (On.Interferon == 0))

### print demographics, MS+noA = 96, MS+mildA = 248, MS+severeA=22
make_demographics_table_ms_anxiety_simple(data_frame = df_demo_and_fascicles_no_steroids_or_ifn[df_demo_and_fascicles_no_steroids_or_ifn$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified",])

make_demographics_table_ms_anxiety_simple(data_frame = df_demo_and_fascicles_no_steroids_or_ifn[((df_demo_and_fascicles_no_steroids_or_ifn$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") & (df_demo_and_fascicles_no_steroids_or_ifn$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds")),])

## Make df for Anxiety Diagnosis
#define group
  uncinate_by_dx_df <- df_demo_and_fascicles_no_steroids_or_ifn %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  ungroup()

 #bilateral UF ~ Diagnosis: DiagnosisMS+Anxiety    0.09339    0.05373   1.738   0.0849 .    trend, likely due to sample size
disease_mean_UF_gam_prop <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_gam_prop)

##test Anxiety Severity Model
#define group
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles_no_steroids_or_ifn %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) #n=372

  #bilat anxiety dose + sex + age: anxiety_dose: anxiety_dose    0.04662    0.02448   1.904   0.0577 .    (trend, likely due to sample size)   
  anxietyGrouping_mean_UF_by_sex_gam_prop <- gam(mean_UF_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam_prop)

   #print summary stats
   gam_summary_stats(disease_mean_UF_gam_prop)
   gam_summary_stats(anxietyGrouping_mean_UF_by_sex_gam_prop)
```

#### Plots 
#####Uncinate R,L,mean Prop
```{r plots_uncinate}
  #proportion injured over R Uncinate prop

#n=372
df_mean_UF_mean_by_anxiety_severity <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) %>%
  mutate(anxiety_severity = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed == "true_healthy", "MS+noA", 
                                   ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed == "Has.Anxietydx.OR.Antianxiety.Meds", "MS+mildA", "MS+severeA"))) %>% 
  mutate(anxiety_severity = factor(anxiety_severity, levels = c("MS+noA", "MS+mildA", "MS+severeA"))) %>%
  mutate(mean_UF_prop = mean_UF_prop*100) %>%
  group_by(anxiety_severity) %>%
  summarise(mean_UF_mean = round(mean(mean_UF_prop , na.rm = TRUE),0),
            sd_UF = sd(mean_UF_prop, na.rm = TRUE),
            sem_UF = sd(mean_UF_prop, na.rm = TRUE) / sqrt(n())) %>%
  mutate(anxiety_severity = factor(anxiety_severity, levels = unique(anxiety_severity))) %>%
  ungroup()

prop_injured_mean_uncinate_by_anxiety_burden_collapsed_vars_plot_bar <-  ggplot(df_mean_UF_mean_by_anxiety_severity, aes(x=anxiety_severity, y=mean_UF_mean, fill=anxiety_severity)) +
  geom_bar(stat="identity", width = .6) +  
  geom_errorbar(aes(ymin = mean_UF_mean - sem_UF, ymax = mean_UF_mean + sem_UF), width = 0.2, size = 0.8)  +
  scale_fill_viridis(discrete = TRUE, option = "D") +  # Use Viridis color scale
  ylim(0,45)+
  theme(axis.text.x = element_text(size=12, colour = "black"),
        axis.text.y = element_text(size=12, colour = "black"),
        axis.title.y = element_text(size=16, colour = "black"),
        axis.title.x = element_blank(),
        axis.line = element_line(linewidth = 1.5, colour="black"),
        legend.position = "none",
        panel.background = element_rect(colour = "white", fill = "white")) + 
  labs(y = "Percent (%) of Uncinate Injured")
prop_injured_mean_uncinate_by_anxiety_burden_collapsed_vars_plot_bar

#save pdf
pdf(file=paste0(homedir, "/results/Figures_for_msanxiety_paper/UF_anxiety_severity_bar.pdf"))
  prop_injured_mean_uncinate_by_anxiety_burden_collapsed_vars_plot_bar
dev.off()

```
##### Uncinate Violin Plot
```{r uncinate_violin}
#### trying prettier things
df_disease_burden_unc_and_anx_net_collapsed_variables <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) %>%
  mutate(anxiety_severity = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed == "true_healthy", "MS+noA", 
                                   ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed == "Has.Anxietydx.OR.Antianxiety.Meds", "MS+mildA", "MS+severeA"))) %>% 
  mutate(anxiety_severity = factor(anxiety_severity, levels = c("MS+noA", "MS+mildA", "MS+severeA"))) %>%
  mutate(mean_UF_prop = mean_UF_prop*100)

# Create the plot with narrower violins and error bars
prop_injured_mean_uncinate_by_anxiety_burden_collapsed_vars_plot <- ggplot(df_disease_burden_unc_and_anx_net_collapsed_variables, aes(x = anxiety_severity, y = mean_UF_prop, fill = anxiety_severity)) +
  geom_violin(width = 0.6, alpha =1.5, color = "black", trim = FALSE) +  # Adjust violin width and transparency
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  ylim(0,100) +
  #geom_errorbar(aes(ymin = mean_UF_prop - sem_UF_prop, ymax = mean_UF_prop + sem_UF_prop), width = 0.2, color = "black") +  # Add error bars
  scale_fill_viridis(discrete = TRUE, option = "D") +  # Use Viridis color scale
  #theme_minimal(base_size = 14) +  # Use a minimal theme with a base font size
  theme(
    axis.text.x = element_text(size = 12, colour = "black"),  # Rotate x-axis labels for better readability
    axis.text.y = element_text(size = 12, colour = "black"),
    axis.title.y = element_text(size = 16, colour = "black"),
    axis.title.x = element_blank(),
    axis.line = element_line(linewidth = 1.5, colour = "black"),
    legend.position = "none",
    panel.background = element_rect(colour = "white", fill = "white"),
    plot.margin = margin(10, 10, 10, 10)  # Add margins around the plot
  ) + 
  labs(y = "Percent (%) of Uncinate Injured")

prop_injured_mean_uncinate_by_anxiety_burden_collapsed_vars_plot

#save pdf
pdf(file=paste0(homedir, "/results/Figures_for_msanxiety_paper/UF_anxiety_severity_violin.pdf"))
  prop_injured_mean_uncinate_by_anxiety_burden_collapsed_vars_plot
dev.off()

```
##### Geom point
```{r geom_point}
#### trying prettier things
df_disease_burden_unc_and_anx_net_collapsed_variables <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) %>%
  mutate(anxiety_severity = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed == "true_healthy", "MS+noA", 
                                   ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed == "Has.Anxietydx.OR.Antianxiety.Meds", "MS+mildA", "MS+severeA"))) %>% 
  mutate(anxiety_severity = factor(anxiety_severity, levels = c("MS+noA", "MS+mildA", "MS+severeA"))) %>%
  mutate(mean_UF_prop = mean_UF_prop*100)

# Create the plot with narrower violins, viridis colors, and horizontal lines at means
prop_injured_mean_uncinate_by_anxiety_burden_collapsed_vars_plot <- ggplot(df_disease_burden_unc_and_anx_net_collapsed_variables, aes(x = anxiety_severity, y = mean_UF_prop, color = anxiety_severity)) +
  geom_point(position = position_jitter(.1), size = 2, alpha = 0.8) +  # Add points with jitter using viridis colors
  stat_summary(fun = "mean", geom = "crossbar", width = 0.3, color = "black", size = 0.5) +  # Add horizontal lines at means
  scale_color_viridis(discrete = TRUE, option = "D") +  # Use Viridis color scale for points
  theme(
    axis.text.x = element_text(size = 12, colour = "black"),  # X-axis labels
    axis.text.y = element_text(size = 12, colour = "black"),  # Y-axis labels
    axis.title.y = element_text(size = 16, colour = "black"),  # Y-axis title
    axis.title.x = element_blank(),  # Remove X-axis title
    axis.line = element_line(linewidth = 1.5, colour = "black"),  # Axis lines
    legend.position = "none",  # Remove legend
    panel.background = element_rect(colour = "white", fill = "white")  # Background
  ) + 
  labs(y = "Percent (%) of Uncinate Injured")  # Y-axis label

# Display the plot
prop_injured_mean_uncinate_by_anxiety_burden_collapsed_vars_plot

#save pdf
pdf(file=paste0(homedir, "/results/Figures_for_msanxiety_paper/UF_anxiety_severity_dot.pdf"))
  prop_injured_mean_uncinate_by_anxiety_burden_collapsed_vars_plot
dev.off()

```

### Sex Differences lm and chisq - N.S.
```{r diseaseXsexdiff}
############## sex diff
#df_anxiety_disorder_by_sex <-

#define group
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter (anxiety_dose != -1) #n=372

###### Sex differences - NONE. important because anxiety dx usually have differences

## in anxiety diagnosis

#####FIRST, SEX DIFF IN FREQUENCY OF ILLNESS - no diff between groups

#first number is # with anxiety, second is number without anxiety
anxiety_female_0 <-df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(anxiety_dose == 0 & osex == "Female") %>% count()
anxiety_total_0 <-df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(anxiety_dose == 0) %>% count()
anxiety_female_1 <-df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(anxiety_dose == 1 & osex == "Female") %>% count()
anxiety_total_1 <-df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(anxiety_dose == 1) %>% count()
anxiety_female_2 <- df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(anxiety_dose == 2 & osex == "Female") %>% count()
anxiety_total_2 <- df_demo_and_fascicles_no_unclass_anxiety_dose %>% filter(anxiety_dose == 2) %>% count()

#make table
num_w_anxiety_female <- as.numeric(c(anxiety_female_0, anxiety_female_1, anxiety_female_2))
num_total <-as.numeric(c(anxiety_total_0,anxiety_total_1, anxiety_total_2))
df_2x3_anxiety_score_by_total_people_per_diagnosis <- data.frame(cbind(num_w_anxiety_female, num_total), row.names = c("MS+noAnxiety","MS+AnxietyorMed", "MS+AnxietyandMed"))

#X-squared (Pearson's) X-squared = 0.23182, df = 2, p-value = 0.8906
chisq_anxiety_dose <- chisq.test(df_2x3_anxiety_score_by_total_people_per_diagnosis)
chisq_anxiety_dose

#just healthy and super anxious/sick, X-squared = 2.9757e-31, df = 1, p-value = 1
num_w_anxiety_female_just_healthy_and_sick <- as.numeric(c(anxiety_female_0, anxiety_female_2))
num_total_just_healthy_and_sick <-as.numeric(c(anxiety_total_0,anxiety_total_2))
df_2x2_anxiety_score_by_total_people_per_diagnosis_num_total_just_healthy_and_sick <- data.frame(cbind(num_w_anxiety_female_just_healthy_and_sick, num_total_just_healthy_and_sick), row.names = c("MS+noAnxiety", "MS+AnxietyandMed"))

chisq_healthy_vs_sick <- chisq.test(df_2x2_anxiety_score_by_total_people_per_diagnosis_num_total_just_healthy_and_sick)
chisq_healthy_vs_sick

#frequency_diff_between_sexes <- chisq.test(df_demo_and_fascicles$Has.Anxietydx.AND.Antianxietymed, df_demo_and_fascicles$osex)

#simple lmer across whole network

#overall - osex.L          -0.0154963  0.0180285  -0.860    0.391 - NS
disease_between_sexes <- lm(proportion_volume_lost_per_total_network_size_sum ~ osex+ PAT_AGE_AT_EXAM, data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(disease_between_sexes)

# mean uncinate L - osex.L          -0.019879   0.022675  -0.877    0.381  ns
disease_between_sexes_UF_mean_prop <- lm(mean_UF_prop ~ osex+ PAT_AGE_AT_EXAM, data = df_demo_and_fascicles_no_unclass_anxiety_dose)
summary(disease_between_sexes_UF_mean_prop)
```

### Plots

##### Overall disease in brain
```{r}
#n = 372

df_disease_burden_unc_and_anx_net_collapsed_variables <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified")

prop_injured_by_anxiety_burden_collapsed_vars_plot <-  ggplot(df_disease_burden_unc_and_anx_net_collapsed_variables, aes(x= oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed , y=proportion_volume_lost_per_total_network_size_sum, fill= oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed )) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  theme(axis.text.x = element_text(size=8, angle = 45, vjust = 0.5))
prop_injured_by_anxiety_burden_collapsed_vars_plot
```

###Promis data in UF - groups too small
#### Uncinate by anxiety diagnosis (N.S) 
```{r uncinate_by_dx}

### Results:  
#define group
  uncinate_by_dx_df <- df_people_with_promis_data_scores_added_unique %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  ungroup() #n=19

#bilateral UF prop ~ Diagnosis: DiagnosisMS+Anxiety -0.02490    0.11510  -0.216   0.8317 
disease_mean_UF_prop_gam <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_prop_gam)

    #print summary stats
 gam_summary_stats(disease_mean_UF_prop_gam)

```

#### Uncinate by Anxiety Dose (N.S.)
```{r uncinateByAnxietyDose}

#define group
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_people_with_promis_data_scores_added_unique %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) #n=64

  #bilat anxiety dose -anxiety_dose   -0.006261   0.055025  -0.114 0.909793       NS
  anxietyGrouping_mean_UF_by_sex_gam_prop <- gam(mean_UF_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam_prop)

    #print summary stats

 gam_summary_stats(anxietyGrouping_mean_UF_by_sex_gam_prop)

```